---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# SAWpaper: Self-Assessment Works Paper
<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/cstatops)](https://CRAN.R-project.org/package=SAWpaper)
<!-- badges: end -->

This package and corresponding GitHub repository are intended to enhance the
reproducibility of a research paper. The principal investigators for the study 
are Dr. Paula Winke (Associate Professor, Department of Linguistics & Germanic, 
Slavic, Asian, and African Languages, Michigan State University) and Dr. Susan 
Gass (University Distinguished Professor, Department of Linguistics & Germanic, 
Slavic, Asian, and African Languages, Michigan State University)  The paper 
analyzes data on Spanish language learners who took a Can-Do self-assessment
test, along with the more authoritative OPIc language proficiency test. We did
both correlation analyses and continuation-ratio models that examine the effect
of course and OPIc speaking proficiency scores on the passing rate for each
level of the Can-Do statements self-assessment. The objective was to validate
the Can-Do test results.

## Task List
* Add disclaimer text regarding our results not being the official position of 
  the funder? 
* Add more details about downloading and installing the repository. 

## Associated Research Paper
Winke, P., Zhang, X., & Pierce, S. J. (2019). *Self-assessment works! 
   Continuation-ratio models for testing course and OPIc score effects on oral
   proficiency self-assessments*. Michigan State University, East Lansing, MI.
   Manuscript in preparation.
   
## Funding Sources 
Funding for this research was provided by the following grants. 

Winke, P., Gass, S., & Pierce, S. J. (08/01/2017-07/31/2018). Michigan State
   University Language Flagship Proficiency Initiative Year 4 continuation 
   proposal. [Award #: 0054-MSU-22-PI-280-PO2, $249,865]. Sponsor: Institute
   of International Education. Location: Michigan State University, East 
   Lansing, MI. 

Winke, P., Gass, S., & Pierce, S. J. (01/01/2019–12/31/2019). Michigan State
   University Language Flagship Proficiency Initiative Year 5 continuation
   proposal. [Award #: 0054-MSU-22-PI-280-PO2, Awarded, $37,634]. Sponsor:
   Institute of International Education. Location: Michigan State University, 
   East Lansing, MI. 

## Software Environment
We use R Markdown to enhance reproducibility. Knitting an R Markdown script
can generate a PDF file containing explanatory text, R code, plus R output 
(text and graphics), although it can also produce HTML or Markdown files too
depending on the contents of the R Markdown file. 

* We use [RStudio (version 1.2.5001 or later)](www.rstudio.org) to work with 
  R and R markdown files. The software chain looks like this:
  **Rmd file -> RStudio -> R -> knitr -> pandoc -> MiKTeX -> PDF file**.
* We are using [MiKTeX version 2.9](https://miktex.org) to compile LaTeX files 
  into PDF files. 

## Installation

This package is only available from a *private* repository available on 
[GitHub](https://github.com/). It is private for now because we are not ready to
release this code yet. 

The package can be installed with the code shown below. If you don't have
*devtools* installed, uncomment that line first. 

```r
# install.packages("devtools")
devtools::install_github("sjpierce/SAWpaper")
```

### Downloading The Repository from GitHub


### Installing the Repository from a Compressed File


## Repository Contents 
This repository contains data files, the R markdown scripts used to analyze 
them, plus the output from when we ran the scripts. There is considerably more
raw statistical output than we could put directly into the resulting manuscript.
That is preserved here for the sake of reproducibility and completeness of 
reporting. We also include other files necessary to work with the repository in 
RStudio and GitHub and to build the R package from the repository files. 

## Data File Documentation
There are some data files included in this repository. The are located in the
*data* folder and stored as comma separated values in text files (i.e., .csv 
files). We briefly describe the contents of each file below. 

```{r example}
# Load some useful packages.
library(here)
library(knitr)

# Read in the included data files. 
df_corr <- read.csv(here::here("./data/DF_CORR.csv"))
df_crm  <- read.csv(here::here("./data/DF_CRM.csv"))
```

### Learner Data File: *data/DF_CORR.csv*
This data file has one row per Spanish language learner in our study, with
$N = `r nrow(df_corr)`$ observations and `r ncol(df_corr)` variables. The first 
few rows of data are shown below. 

```{r}
# Display top few rows of df_corr.
kable(head(df_corr))
```

This is the data file to use when examining simple correlations between
variables like OPIC and Course. The variables are described below. Remember that 
R is case-sensitive with variable names, so OPIc and OPIC are different (but
related) variables.

- *id*: An anonymized participant identification number. Each learner has a unique 
  id value. The values are integer numbers. 
- *Course*: The level of the Spanish language course the learner was taking at 
  the time of taking the self-assessment test. First year courses are coded 100,
  second year courses are coded 200, third year courses are coded 300, and 
  fourth year courses are coded 400. Course values are stored as integers. 
- *OPIc*: These two-character alphabetic codes are values for scores on the 
  American Council on the Teaching of Foreign Languages (ACTFL) Oral Proficiency
  Interview - Computer (OPIc) test for Spanish language proficiency. The 
  possible scores, in increasing level of proficiency, are: 

    - *NL*: Novice low.
    - *NM*: Novice mid.
    - *NH*: Novice high.
    - *IL*: Intermediate low.
    - *IM*: Intermediate mid.
    - *IH*: Intermediate high.
    - *AL*: Advanced low.
    - *AM*: Advanced mid.
    - *AH*: Advanced high.
    - *S*: Superior (which does not occur in thie dataset). 

- *OPIC*: These integer values are a simple recode of the OPIc variable. The 
  possible scores, in increasing level of proficiency, are: 

    - *1*: Novice low.
    - *2*: Novice mid.
    - *3*: Novice high.
    - *4*: Intermediate low.
    - *5*: Intermediate mid.
    - *6*: Intermediate high.
    - *7*: Advanced low.
    - *8*: Advanced mid.
    - *9*: Advanced high.
    - *10*: Superior (which does not occur in thie dataset). 

- *Level*: These integer values show the maximum level the learner reached on
  the self-assessment test based on 50 Can-Do statements. There are five 
  possible levels (1 through 5), with higher values indicating greater 
  self-assessed proficiency. 

### Level-Transition Test Data File: *data/DF_CRM.csv*
This data file has one row per level transition test attempted by a learner in
our study, with $N = `r nrow(df_crm)`$ observations and `r ncol(df_crm)`
variables. The first few rows of data are shown below.

```{r}
# Display top few rows of df_crm.
kable(head(df_crm))
```

This dataset is used in continuation-ratio modeling of the self-assessment 
test results. The data shown above have two rows for id = 3. They show that this
learner passed the level transition test to get from level 1 to level 2, but
failed to pass the transition test to get from level 2 to level 3. The variables
are described in more detail below. 

- *id*: An anonymized participant identification number. Each learner has a unique 
  id value. The values are integer numbers. 
- *Course*: The level of the Spanish language course the learner was taking at 
  the time of taking the self-assessment test. First year courses are coded 100,
  second year courses are coded 200, third year courses are coded 300, and 
  fourth year courses are coded 400. Course values are stored as integers. 
- *COPIC*: These integer values are a centered version of the OPIC variable 
  found in *df_corr*. The centering was done around OPIC = 5, so the centered
  scores, in increasing level of proficiency, are: 

    - *-4*: Novice low.
    - *-3*: Novice mid.
    - *-2*: Novice high.
    - *-1*: Intermediate low.
    - *0*: Intermediate mid.
    - *1*: Intermediate high.
    - *2*: Advanced low.
    - *3*: Advanced mid.
    - *4*: Advanced high.
    - *5*: Superior (which does not occur in thie dataset). 

- *Level*: These integer values show which level transition test from the Can-Do
  self-assessment the learner was attempting pass on a given row of data. There
  are four possible level transition tests: 

    - *1*: Transition from level 1 to level 2.
    - *2*: Transition from level 2 to level 3.
    - *3*: Transition from level 3 to level 4.
    - *4*: Transition from level 4 to level 5.
  
- *pass*: This binary integer variable is coded 0 when the learner did not pass
  the level transition test shown in the Level variable, and 1 when the learner 
  pased the test. 
