---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# SAWpaper: Self-Assessment Works Paper Research Compendium
<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/cstatops)](https://CRAN.R-project.org/package=SAWpaper)
<!-- badges: end -->

This package and corresponding GitHub repository are intended to enhance the
reproducibility of a research paper by serving as a research compendium
(Marwick, Boettiger, & Mullen, 2018). The principal investigators for the study
are Dr. Paula Winke (Professor) and Dr. Susan Gass (University Distinguished 
Professor), both of whom belong to the Department of Linguistics, Languages, & 
Cultures at Michigan State University. The paper analyzes data on Spanish 
language learners who took a Can-Do self-assessment test, along with the more 
authoritative OPIc language proficiency test. We did both correlation analyses 
and continuation-ratio models that examine the effect of course and OPIc 
speaking proficiency scores on the passing rate for each level of the Can-Do 
statements self-assessment. The objective was to validate the Can-Do test 
results.

## Task List
* Add disclaimer text regarding our results not being the official position of 
  the funder? 
* Add more details about downloading and installing the repository. 
* Update the data documentation to point toward the published data archive at 
  ICPSR. 
* Update `inst/SAW_Paper_Spanish_Analysis.Rmd` to integrate code from Xiaowan's
  file `CRM_Spanish_2020-2-3.Rmd`. 
* Copy Xiaowan's custom functions from `CRM_Spanish_2020-2-3.Rmd` into R scripts
  under `R/` folder and add roxygen documentation to generate help file.
* Fix check() warning about Undocumented data sets: 'df_corr' 'df_crm'. 
   All user-level objects in a package should have documentation entries. See
   chapter 'Writing R documentation files' in the 'Writing R Extensions' manual.

## Associated Research Paper
Winke, P., Zhang, X., & Pierce, S. J. (2021). 
*A closer look at a marginalized test method: Self-assessment as a measure of speaking proficiency* 
[Manuscript submitted for publication]. Department of Linguistics, Languages, & 
Cultures, Michigan State University. 

## Funding Sources 
Funding for this research was provided by the following grants. 

Winke, P., Gass, S., & Pierce, S. J. (08/01/2017–12/31/2018). 
*Michigan State University Language Flagship Proficiency Initiative Year 4 continuation proposal* 
(Award #: 0054-MSU-22-PI-280-PO2) [Grant]. Institute of International Education. 

Winke, P., Gass, S., & Pierce, S. J. (01/01/2019–12/31/2019). 
*Michigan State University Language Flagship Proficiency Initiative Year 5 continuation proposal* 
(Award #: 0054-MSU-22-PI-280-PO2) [Grant]. Institute of International Education. 


## Software Environment
We use R Markdown to enhance reproducibility because it provides excellent
support for generating dynamic reports (Mair, 2016). Knitting an R Markdown
script can generate a PDF output file containing explanatory text, R code, plus
R output (text and graphics), although it can also produce HTML or Markdown
files too depending on the contents of the R Markdown file.

- We use [RStudio](www.rstudio.org) to work with R and R Markdown files. The 
  software chain looks like this:
  **Rmd file -> RStudio -> R -> knitr -> pandoc -> TinyTex -> PDF file**. 
- A version of [pandoc](https://pandoc.org/) comes bundled with RStudio, but if 
  you want the most recent version, download it from https://pandoc.org/. 
- We use Git (Torvalds et al., 2020) for version control, with the primary 
  (i.e., master) repository hosted online by [GitHub](https://github.com/). 
  For a short introduction to Git, see Bryan (2018). There is a longer, more
  detailed resource on using Git with R as well (Bryan et al., 2019). 
- We recommend using [TinyTeX](https://yihui.org/tinytex/) to compile LaTeX 
  files produced by pandoc into PDF files. However, it may be viable to use 
  [MiKTeX](https://miktex.org) instead. 
- This package depends on functions from the *piercer* package, which is 
  available from a public GitHub repository at 
  https://github.com/sjpierce/piercer. Please read and follow the installation
  instructions for *piercer* before trying to use this package. 

## Installation
This package is only available from a *private* repository available on 
[GitHub](https://github.com/) at at https://github.com/sjpierce/SAWpaper. It is
private for now because we are not ready to release this code yet. The package 
can be installed with the code shown below. If you don't have
*devtools* installed, uncomment that line first. 

```r
# install.packages("devtools")
devtools::install_github("sjpierce/SAWpaper")
```

### Obtain the Repository from GitHub
If you use Git (Torvalds et al., 2020) and have a GitHub account, either clone
or fork and clone the package to your computer using the usual Git commands
(Bryan, 2018; Bryan et al., 2019, Chapters 28 and 32). Otherwise, manually
download a ZIP file from https://github.com/sjpierce/SAWpaper and unzip its
contents to a folder. Either way, the files should end up in a folder called
*SAWpaper* on your computer. That folder is your local copy of the repository.

Note that the package code uses relative rather than absolute folder path and
file name references. Moving or renaming subfolders and/or files may cause
problems. We have tested it only with the folder structure and file naming used
in the primary repository on GitHub.

## Repository Contents 
This repository contains data files, the R markdown scripts used to analyze 
them, plus the output from when we ran the scripts. There is considerably more
raw statistical output than we could put directly into the resulting manuscript.
That is preserved here for the sake of reproducibility and completeness of 
reporting. We also include other files necessary to work with the repository in 
RStudio and GitHub and to build the R package from the repository files. 

## Data File Documentation
There are some data files included in this repository. The are located in the
*data* folder and stored as comma separated values in text files (i.e., .csv 
files). We briefly describe the contents of each file below. 

```{r example}
# Load some useful packages.
library(here)
library(knitr)

# Read in the included data files. 
df_corr <- read.csv(here::here("./data/DF_CORR.csv"))
df_crm  <- read.csv(here::here("./data/DF_CRM.csv"))
```

### Learner Data File: *data/DF_CORR.csv*
This data file has one row per Spanish language learner in our study, with
$N = `r nrow(df_corr)`$ observations and `r ncol(df_corr)` variables. The first 
few rows of data are shown below. 

```{r}
# Display top few rows of df_corr.
kable(head(df_corr))
```

This is the data file to use when examining simple correlations between
variables like OPIC and Course. The variables are described below. Remember that 
R is case-sensitive with variable names, so OPIc and OPIC are different (but
related) variables.

- *id*: An anonymized participant identification number. Each learner has a unique 
  id value. The values are integer numbers. 
- *Course*: The level of the Spanish language course the learner was taking at 
  the time of taking the self-assessment test. First year courses are coded 100,
  second year courses are coded 200, third year courses are coded 300, and 
  fourth year courses are coded 400. Course values are stored as integers. 
- *OPIc*: These two-character alphabetic codes are values for scores on the 
  American Council on the Teaching of Foreign Languages (ACTFL) Oral Proficiency
  Interview - Computer (OPIc) test for Spanish language proficiency. The 
  possible scores, in increasing level of proficiency, are: 

    - *NL*: Novice low.
    - *NM*: Novice mid.
    - *NH*: Novice high.
    - *IL*: Intermediate low.
    - *IM*: Intermediate mid.
    - *IH*: Intermediate high.
    - *AL*: Advanced low.
    - *AM*: Advanced mid.
    - *AH*: Advanced high.
    - *S*: Superior (which does not occur in thie dataset). 

- *OPIC*: These integer values are a simple recode of the OPIc variable. The 
  possible scores, in increasing level of proficiency, are: 

    - *1*: Novice low.
    - *2*: Novice mid.
    - *3*: Novice high.
    - *4*: Intermediate low.
    - *5*: Intermediate mid.
    - *6*: Intermediate high.
    - *7*: Advanced low.
    - *8*: Advanced mid.
    - *9*: Advanced high.
    - *10*: Superior (which does not occur in thie dataset). 

- *Level*: These integer values show the maximum level the learner reached on
  the self-assessment test based on 50 Can-Do statements. There are five 
  possible levels (1 through 5), with higher values indicating greater 
  self-assessed proficiency. 

### Level-Transition Test Data File: *data/DF_CRM.csv*
This data file has one row per level transition test attempted by a learner in
our study, with $N = `r nrow(df_crm)`$ observations and `r ncol(df_crm)`
variables. The first few rows of data are shown below.

```{r}
# Display top few rows of df_crm.
kable(head(df_crm))
```

This dataset is used in continuation-ratio modeling of the self-assessment 
test results. The data shown above have two rows for id = 3. They show that this
learner passed the level transition test to get from level 1 to level 2, but
failed to pass the transition test to get from level 2 to level 3. The variables
are described in more detail below. 

- *id*: An anonymized participant identification number. Each learner has a unique 
  id value. The values are integer numbers. 
- *Course*: The level of the Spanish language course the learner was taking at 
  the time of taking the self-assessment test. First year courses are coded 100,
  second year courses are coded 200, third year courses are coded 300, and 
  fourth year courses are coded 400. Course values are stored as integers. 
- *COPIC*: These integer values are a centered version of the OPIC variable 
  found in *df_corr*. We centered around Intermediate-mid (OPIC = 5) because it
  is the level of proficiency expected of students exiting the 200-level
  courses. The centered scores, in increasing level of proficiency, are:

    - *-4*: Novice low.
    - *-3*: Novice mid.
    - *-2*: Novice high.
    - *-1*: Intermediate low.
    - *0*: Intermediate mid (expected proficiency for students after 200-level 
      courses).
    - *1*: Intermediate high.
    - *2*: Advanced low.
    - *3*: Advanced mid.
    - *4*: Advanced high.
    - *5*: Superior (which does not occur in thie dataset). 

- *Level*: These integer values show which level transition test from the Can-Do
  self-assessment the learner was attempting pass on a given row of data. There
  are four possible level transition tests: 

    - *1*: Transition from level 1 to level 2.
    - *2*: Transition from level 2 to level 3.
    - *3*: Transition from level 3 to level 4.
    - *4*: Transition from level 4 to level 5.
  
- *pass*: This binary integer variable is coded 0 when the learner did not pass
  the level transition test shown in the Level variable, and 1 when the learner 
  pased the test. 

# References
Bryan, J. (2018). Excuse me, do you have a moment to talk about version control? 
*The American Statistician, 72*(1), 20-27. 
[doi:10.1080/00031305.2017.1399928](https://doi.org/10.1080/00031305.2017.1399928)

Bryan, J., The STAT 545 TAs, & Hester, J. (2019). Happy Git and GitHub for the 
useR. Retrieved from https://happygitwithr.com

Mair, P. (2016). Thou shalt be reproducible! A technology perspective.
*Frontiers in Psychology, 7*(1079), 1-17. 
[doi:10.3389/fpsyg.2016.01079](http://dx.doi.org/10.3389/fpsyg.2016.01079)

Marwick, B., Boettiger, C., & Mullen, L. (2018). Packaging data analytical work 
reproducibly using R (and friends). *The American Statistician, 72*(1), 80-88.
[doi:10.1080/00031305.2017.1375986](https://doi.org/10.1080/00031305.2017.1375986)

Torvalds, L., Hamano, J. C., & other contributors to the Git Project. (2020).
Git for Windows (Version 2.26.2) [Computer program]. Brooklyn, NY: Software 
Freedom Conservancy. Retrieved from https://git-scm.com
