---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# SAWpaper: Self-Assessment Works Paper
<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/cstatops)](https://CRAN.R-project.org/package=SAWpaper)
<!-- badges: end -->

This package and corresponding GitHub repository are intended to enhance the
reproducibility of a research paper. The principal investigators for the study 
are Dr. Paula Winke (Associate Professor, Department of Linguistics & Germanic, 
Slavic, Asian, and African Languages, Michigan State University) and Dr. Susan 
Gass (University Distinguished Professor, Department of Linguistics & Germanic, 
Slavic, Asian, and African Languages, Michigan State University)  The paper 
analyzes data on Spanish language learners who took a Can-Do self-assessment
test, along with the more authoritative OPIc language proficiency test. We did
both correlation analyses and continuation-ratio models that examine the effect
of course and OPIc speaking proficiency scores on the passing rate for each
level of the Can-Do statements self-assessment. The objective was to validate
the Can-Do test results.

## Installation

This package is only available from a *private* repository available on 
[GitHub](https://github.com/). It is private for now because we are not ready to
release this code yet. 

The package can be installed with the code shown below. If you don't have
*devtools* installed, uncomment that line first. 

```r
# install.packages("devtools")
devtools::install_github("sjpierce/SAWpaper")
```

## Heading
Put text here.  

## Example
Text goes here. 

```{r example}
#library(SAWpaper)
# Need to add example code
```

## Task List
* Improve this file. 
* Create a script to generate a data flow diagram. 
