---
title: \sffamily{\bfseries{\LARGE Self-Assessment Works! Import Data}}
geometry: "left=1.0in,right=1.0in,top=.75in,bottom=.75in"
author: "Steven J. Pierce & Xiaowan Zhang"
output:
 pdf_document:
   latex_engine: xelatex
   number_sections: true
   toc: yes
   toc_depth: 3
urlcolor: blue
header-includes:
- \usepackage{fancyhdr}
- \usepackage[yyyymmdd,hhmmss]{datetime}
- \usepackage{lastpage}
- \usepackage{amsmath,amsthm}
- \usepackage{unicode-math}
- \defaultfontfeatures{Ligatures=TeX}
- \usepackage[font={small}, margin=1cm, skip=2pt]{caption}
- \usepackage{url}
- \usepackage{floatrow} 
- \floatplacement{figure}{!ht}
- \floatplacement{table}{!ht}
- \usepackage{placeins}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{dcolumn}
- \usepackage{titling}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \pretitle{\begin{center}\LARGE\bfseries}
- \posttitle{\end{center}}
- \usepackage{lscape}
- \pagestyle{fancy}
- \lhead{Self-Assessment Works! (Spanish) Import Data}
- \rhead{\today\ \currenttime}
- \cfoot{ }
- \lfoot{\texttt{\footnotesize \detokenize{`r sub(".Rmd", ".pdf", knitr:::current_input(dir = FALSE))`}}} 
- \fancyfoot[R]{\thepage\ of \pageref*{LastPage}}
- \renewcommand{\headrulewidth}{0.4pt}
- \renewcommand{\footrulewidth}{0.4pt}
- \fancypagestyle{plain}{\pagestyle{fancy}}
---

\FloatBarrier

# Purpose
This file facilitates reproducing the results reported in our manuscript (Winke,
Zhang, & Pierce, 2019), which was based on a presentation (Winke, Pierce, &
Zhang 2018). It imports data on Spanish language students from SPSS data files 
into R and saves out .RData files for use by subsequent scripts. 

# Setup
Set global R chunk options (local chunk options will over-ride global options). 
The method for creating a size option that controls font size in code chunks and
their text output is based on an answer to a question posted on   
[stackoverflow.com](https://stackoverflow.com/questions/25646333/code-chunk-font-size-in-rmarkdown-with-knitr-and-latex/46526740#46526740). 

``` {r global-options, cfsize = "footnotesize"}
# Create a custom chunk hook/option for controlling font size in chunk & output.
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$cfsize != "normalsize", paste0("\n \\", options$cfsize,"\n\n", 
                                              x, "\n\n \\normalsize"), x)
  })

# Global chunk options (over-ridden by local chunk options)
knitr::opts_chunk$set(include  = TRUE, echo = TRUE, error = TRUE, 
                      message = TRUE, warning = TRUE, fig.pos = "!ht", 
                      cfsize = "footnotesize")

# Declare location of this script relative to the project root directory.
here::i_am(path = "inst/SAW_Paper_Import_Data.Rmd")
```

Some of the functions we use below come from a personal R package called
[*piercer*](https://github.com/sjpierce/piercer) that is available online from 
[GitHub](https://github.com). Please read and follow the installation 
instructions for *piercer* at https://github.com/sjpierce/piercer
before knitting this script, otherwise the next chunk will not load that 
package and various parts of this script will not work.  

Load contributed R packages that we need to get additional functions. 

``` {r load-packages}
library(here)              # for here()
library(devtools)          # for session_info()
# Set package options. 
# options(knitr.kable.NA = '0.00')
library(rmarkdown)         # for render(), pandoc_version(). 
library(knitr)             # for kable()
library(tidyr)             # for unite()
library(dplyr)             # for filter(), select(), etc.
library(sjlabelled)       # For set_label(), get_label()
library(haven)            # for read_spss()
# Set package options. 
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)        # for kable_styling(), add_header_above(), 
                           # column_spec(), collapse_rows(), and landscape()
library(piercer)           # for git_report(), which_latex()
```

\FloatBarrier

# Import Data
Please see the `README.Rmd` or `README.md` files in this *SAWpaper* package for
details on obtaining the SPSS data files referenced by the code below. 

\FloatBarrier

## Student-Level Data (for Correlation Analyses)
The file `df_corr.sav` has the cleaned self-assessment data on a set of Spanish
students, with one row per student. We prepared an SPSS file to simplify 
archiving and attaching meta-data such as variable and value labels, level of 
measurement, and so on. The chunk below reads that file into a data frame called 
`df_corr`, then converts a couple variables to proper factors. 

```{r import-student-data}
df_corr <- read_sav(here::here("inst/extdata/df_corr.sav"), user_na = TRUE) %>% 
  mutate(Gender = as_factor(Gender),
         Course = as_factor(Course),
         LevelF = as_factor(Level, levels = "values"))
```

There are `r nrow(df_corr)` rows and `r ncol(df_corr)` variables in `df_corr`. 

\FloatBarrier

## Testlet-Level Data (for Continuation-Rato Modeling)
The file `df_crm.sav` contains data about the same set of students that has been
expanded to one row per student per self-assessment testlet attempted. This data
structure is conceptually similar to a person-period file for longitudinal
analysis. The chunk below reads that file into a data frame called `df_crm`, 
then converts a couple variables to proper factors. We need the expanded 
`df_crm` data frame specifically to run continuation-ratio models. Most other 
analyses are performed based on the `df_corr` data frame. 

```{r import-crm-testlet-data}
df_crm <- read_sav(here::here("inst/extdata/df_crm.sav"), user_na = TRUE) %>% 
  mutate(Gender = as_factor(Gender),
         Course = as_factor(Course),
         Testlet = as_factor(Testlet, levels = "values"),
         MaxLevel = as_factor(MaxLevel, levels = "values"))
```

There are `r nrow(df_crm)` rows and `r ncol(df_crm)` variables in `df_crm`. 

\FloatBarrier

### Contrast Coding 
Tables \ref{tab:Testlet-contrasts} and \ref{tab:Course-contrasts} below show the 
default contrast coding associated with key factor variables in `df_crm` so 
readers know how to interpret model coefficients later. The default for factors 
should be dummy coding (treatment contrasts) with the first level set as the
reference group. 

```{r Testlet-contrasts}
# Table caption and footnote text. 
TCap <- "Default Contrast Dummy Coding for Testlet"
FN   <- "The reference value is the one always coded as zero."

# Generate a table showing the default contrast coding. 
contrasts(df_crm$Testlet) %>% 
  kable(format = "latex", booktabs = TRUE, row.names = TRUE, caption = TCap) %>% 
  add_header_above(c("Testlet", "Contrast Coding" = 3)) %>% 
  column_spec(1:3, width = ".5cm") %>% 
  footnote(general = FN, general_title = "Note: ", footnote_as_chunk = TRUE,
           threeparttable = TRUE)
```

```{r Course-contrasts}
# Table caption and footnote text. 
TCap <- "Default Contrast Dummy Coding for Course"
FN   <- "The reference value is the one always coded as zero."

# Generate a table showing the default contrast coding. 
contrasts(df_crm$Course) %>% 
  kable(format = "latex", booktabs = TRUE, row.names = TRUE, caption = TCap) %>% 
  add_header_above(c("Course", "Contrast Coding" = 3)) %>% 
  column_spec(1:3, width = ".5cm") %>% 
  footnote(general = FN, general_title = "Note: ", footnote_as_chunk = TRUE,
           threeparttable = TRUE)
```

### OPIC Scores
We treat OPIc speaking test proficiency scores as a continuous covariate that
ranges from Novice-low to Superior. Table \ref{tab:OPIC-scores} shows the coding
for the `OPICN` and `COPIC` variables in `df_corr`. The former variable contains
the OPIc ratings in their original numeric form and has value labels attached
that show the meaning of scores. We already removed the cases for which the OPIc
ratings were not meaningful for this study (AR, BR, and UR) before archiving the
file. To make model coefficients more interpretable, we computed centered OPIc
scores and stored them as `COPIC`. A score of zero in COPIC represents the
Intermediate-mid level on the OPIc scale.

```{r OPIC-scores}
# Table caption and footnote text. 
TCap <- paste("OPIc Score Labels, Numeric Values, Centered Values, and",
              "Frequency")
FN   <- paste("COPIC is a centered form of OPICN (COPIC = OPICN - 5) used to",
              "make model intercept terms more interpretable.")

data.frame(OPIC_Label = get_labels(df_crm$OPICN), 
           OPICN = c(as.numeric(sort(unique(df_crm$OPICN))), 10), 
           COPIC = c(as.numeric(sort(unique(df_crm$COPIC))), 5),
           Frequency = as.numeric(c(xtabs(~OPICN, data = df_crm), 0))) %>% 
  kable(format = "latex", booktabs = TRUE, 
        col.names = c("Score Label", "OPICN", "COPIC", "Frequency"), 
        caption = TCap) %>% 
  column_spec(1, width = "4cm") %>% 
  column_spec(2:4, width = "1.5cm") %>% 
  footnote(general = FN, general_title = "Note: ", footnote_as_chunk = TRUE,
           threeparttable = TRUE)
```


# Exploratory Data Analysis

\FloatBarrier

## Descriptive Statistics (Student-Level Data)
Here we make some descriptive results reported in the manuscript reproducible. 

```{r}
# Frequency table for maximum level achieved. 
df_corr %>% 
   group_by(LevelF) %>% 
   summarise(N = n(),
             Pct = 100*N/nrow(df_corr)) %>% 
   ungroup() %>% 
   mutate(CumN = cumsum(N),
          CumPct = cumsum(Pct),
          RemainN = sum(N) - CumN,
          RemainPct = 100 - CumPct) ->
   T01

# Extract N and percent of students with self-assessed level > 1 (i.e., >= 2). 
N_SA_GE_2   <- T01$RemainN[1]
Pct_SA_GE_2 <- T01$RemainPct[1]

# Extract N and percent of students with self-assessed level > 2 (i.e., >= 3). 
N_SA_GE_3   <- T01$RemainN[2]
Pct_SA_GE_3 <- T01$RemainPct[2]

# Display formatted table.
kable(T01, format = "latex", booktabs = TRUE, digits = 0, 
      col.names = c("Level", "N", "%", "Cumulative N", "Cumulative %", 
                    "Remaining N", "Remaining %"),
      caption = "Frequency Table for Maximum Self-Assesment Level Achieved")
```

Only $N = `r N_SA_GE_2`$ (`r round(Pct_SA_GE_2, digits = 0)`%) of the students 
reached level 2 or higher on the self-assessment. Only $N = `r N_SA_GE_3`$ 
(`r round(Pct_SA_GE_3, digits = 0)`%) students reached level 3 or higher on the 
self-assessment. 

\FloatBarrier

## Crosstabulations (Student Data)
We now examine the relationships among Level, Course, and OPIC by calling for
crosstabulations. First, we use the data frame df_corr.

```{r}
# Examine the relationship between SA level and course level. 
T02 <- addmargins(xtabs(~LevelF + Course, data = df_corr))
kable(T02, format = "latex", booktabs = TRUE, 
      caption = "Crosstab Self-Assessment Level by Course") %>%
kable_styling(latex_options = c("repeat_header"))
```

\newpage

```{r}
# Examine the relationship between OPIc scores and SA level.
T03a <- addmargins(xtabs(~LevelF + OPICN, data = df_corr))

# Examine the relationship between OPIc scores and course level.
T03b <- addmargins(xtabs(~Course + OPICN, data = df_corr))

# Combine two cross-tabs into a single object. 
T03  <- rbind(T03a, T03b) 

# Display formatted table. 
kable(T03, format = "latex", booktabs = TRUE,  
      caption = "Crosstabs of Self-Assessment Level and Course by OPIc Score") %>% 
kable_styling(latex_options = c("repeat_header")) %>% 
add_header_above(header = c(" ", "OPIc Score" = 9, " ")) %>% 
group_rows(group_label = "SA Level", bold = FALSE, start_row = 1, end_row = 6) %>% 
group_rows(group_label = "Course", bold = FALSE, start_row = 7, end_row = 11) %>% 
row_spec(row = c(6, 11), italic = TRUE) %>% 
column_spec(column = 11, italic = TRUE)
```

\FloatBarrier

## Crosstabulations (Testlet Data)
We then examine some additional crosstabulations based on the data frame df_crm.
Note that Level in df_crm denotes the transitions between SA levels.

```{r}
# Crosstab to get a look at how often people pass each level transition.  
addmargins(xtabs(~Testlet + Pass, data = df_crm)) %>% 
  as_tibble(.) %>% 
  pivot_wider(names_from = Pass, values_from = n) %>% 
  mutate(CPR = case_when(Testlet != "Sum" ~ `1`/Sum),
         UPR = case_when(Testlet != "Sum" ~ `1`/n_distinct(df_crm$StudentID))) ->
  T04

kable(T04, format = "latex", booktabs = TRUE, digits = 3,
      caption = "Crosstab Self-Assessment Level Transition Testlet by Pass") %>%
kable_styling(latex_options = c("repeat_header")) %>% 
add_header_above(header = c(" ", "Pass" = 2, " ", " ", " ")) %>% 
add_header_above(header = c(" ", "Counts" = 3, "Pass Rates" = 2)) %>% 
row_spec(row = 5, italic = TRUE) %>%
column_spec(column = 4, italic = TRUE)
```

```{r}
# Examine the relationship between SA level transition and course level.
T05 <- addmargins(xtabs(~Testlet + Course, data = df_crm))
kable(T05, format = "latex", booktabs = TRUE, 
      caption = "Crosstab Self-Assessment Level Transition Testlet by Course") %>%
kable_styling(latex_options = c("repeat_header")) %>% 
add_header_above(header = c(" ", "Course" = 4, " ")) %>% 
group_rows(group_label = "Testlet", bold = FALSE, start_row = 1, 
           end_row = 5) %>%
row_spec(row = 5, italic = TRUE) %>%
column_spec(column = 6, italic = TRUE)
```

```{r}
# Examine the relationship between centered OPIc scores and SA level transition.
T06 <- addmargins(xtabs(~Testlet + COPIC, data=df_crm)) 
kable(T06, format = "latex", booktabs = TRUE, 
      caption="Crosstab Self-Assessment Level Transition Testlet by OPIc rating") %>%
kable_styling(latex_options = c("repeat_header")) %>% 
add_header_above(header = c(" ", "Centered OPIc Score (OPIc - 5)" = 9, " ")) %>% 
group_rows(group_label = "Testlet", bold = FALSE, start_row = 1, 
           end_row = 5) %>%
row_spec(row = 5, italic = TRUE) %>%
column_spec(column = 11, italic = TRUE)
```

\FloatBarrier

# Save Data 
The chunks below save the `df_corr` and `df_crm` data frames to an .RData file
in the local copy of the repository for use by subsequent scripts. 

```{r save-data}
save(df_corr, df_crm, file = here::here("data/SAW_Paper_Data.RData"))
```

\FloatBarrier

# References
Winke, P., Zhang, X., & Pierce, S. J. (2021). 
*A closer look at a marginalized test method: Self-assessment as a measure of speaking proficiency* 
[Manuscript submitted for publication]. Department of Linguistics, Languages, & 
Cultures, Michigan State University. 

Winke, P., Pierce, S. J., & Zhang, X. (2018, October 12-13). 
*Self-assessment works! Continuation-ratio models for testing course and OPIc score effects on oral proficiency self-assessments* 
[Paper presentation]. East Coast Organization of Language Testers 2018 
Conference, Princeton, NJ, United States. 

\FloatBarrier

# Software Information
We use R Markdown to enhance reproducibility. Knitting the source R Markdown 
script *`r knitr:::current_input()`* generates this PDF file containing 
explanatory text, R code, plus R output (text and graphics).

- We used [RStudio](www.rstudio.org) `r rstudioapi::versionInfo()$version` to 
  work with R and R markdown files. The software chain looks like this:
  **Rmd file > RStudio > R > rmarkdown > knitr > pandoc > TinyTeX or MiKTeX > PDF file**.
- We recommend using [TinyTeX](https://yihui.org/tinytex/) to compile LaTeX 
  files into PDF files. However, it should be viable to use 
  [MiKTeX](https://miktex.org) instead. 
- We used [pandoc](https://pandoc.org) `r pandoc_version()` for this 
  document. 

This document was generated using the following computational environment and 
dependencies: 

``` {r show_citations}
# Check and report whether we used TinyTex or other LaTeX software. 
which_latex()

# Get R and R package version numbers in use.
devtools::session_info()
```

The current Git commit details and status are:

```{r git_details, echo = TRUE, eval=TRUE}
git_report()
```
