---
title: \sffamily{\bfseries{\LARGE Self-Assessment Works! Paper Analyze Data}}
geometry: "left=1.0in,right=1.0in,top=.75in,bottom=.75in"
author: "Steven J. Pierce & Xiaowan Zhang"
output:
 pdf_document:
   latex_engine: xelatex
   number_sections: true
   toc: yes
   toc_depth: 3
urlcolor: blue
header-includes:
- \usepackage{fancyhdr}
- \usepackage[yyyymmdd,hhmmss]{datetime}
- \usepackage{lastpage}
- \usepackage{amsmath,amsthm}
- \usepackage{unicode-math}
- \defaultfontfeatures{Ligatures=TeX}
- \usepackage[font={small}, margin=1cm, skip=2pt]{caption}
- \usepackage{url}
- \usepackage{floatrow} 
- \floatplacement{figure}{!ht}
- \floatplacement{table}{!ht}
- \usepackage{placeins}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{dcolumn}
- \usepackage{titling}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage{lscape}
- \pagestyle{fancy}
- \lhead{Self-Assessment Works! Analyze Data}
- \rhead{\today\ \currenttime}
- \cfoot{ }
- \lfoot{\texttt{\footnotesize \detokenize{`r sub(".Rmd", ".pdf", knitr:::current_input(dir = FALSE))`}}} 
- \fancyfoot[R]{\thepage}
- \renewcommand{\headrulewidth}{0.4pt}
- \renewcommand{\footrulewidth}{0.4pt}
- \fancypagestyle{plain}{\pagestyle{fancy}}
---

\FloatBarrier

# Purpose
This file reproduces the results reported in our manuscript (Winke, Zhang, &
Pierce, 2022), which was based on a presentation (Winke, Pierce, & Zhang 2018).
It analyzes data on Spanish language learners who took a Can-Do self-assessment
test, along with the more authoritative OPIc language proficiency test. We did
both correlation analyses and continuation-ratio models that examine the effect
of course and OPIc speaking proficiency scores on the passing rate for each
level of the Can-Do statements self-assessment. The objective was to validate
the Can-Do test results.

\FloatBarrier

## Target Journal
We submitted this as a "Research Report" to a journal called 
Studies in Second Language Acquisition (SSLA), 
https://www.cambridge.org/core/journals/studies-in-second-language-acquisition.
The author instructions are at
https://www.cambridge.org/core/journals/studies-in-second-language-acquisition/information/instructions-contributors.

We will consider applying for the SSLA *Open Data Badge* and the 
*Open Materials Badge* described at 
https://www.cambridge.org/core/journals/studies-in-second-language-acquisition/open-science-badges
. Hosting a public Github repository should meet the requirements for 
both badges because GitHub is on the 
[Registry of Research Data Repositories](https://www.re3data.org). However, we
may put the actual data in another public repository and just use GitHub for the
code. 

\FloatBarrier

## Comment on Statistical Methodology
We experimented with applying ideas discussed in Wasserstein, Schirm, and Lazar
(2019) about abandoning declarations of statistical significance. This landmark
editorial paper produced by the American Statistical Association discussed
various ways to supplement p-values with additional statistics or replace them 
altogether. We computed some supplemental statistics and include them in this 
set of materials but omitted them from the manuscript. 

We computed Shannon information values (s-values; Greenland, 2019; Wasserstein,
Schirm, & Lazar, 2019). S-values are a rescaling of p-values, such that 
$s = -log_2(p)$. They can range from $s = 0$ when $p = 1$ to $s = \infty$ when 
$p = 0$. Larger s-values correspond to greater evidence against the null 
hypothesis. An s-value can be interpreted as how many bits of information there
are against the hypothesis. To make that easier to understand, suppose you want
to flip a coin repeatedly to determine whether it is a fair coin rather than one
biased toward landing on heads. Each flip provides one bit of information, but
only those resulting in heads are information against the null hypothesis. A
fair coin should yield heads half of the time and tails the other half
(independent with probability = .50 for each outcome on each flip). The null
hypothesis is that the coin is fair. Seeing 2 consecutive heads come up in a set
of 2 flips ($p = 0.5^{2} = 0.25, s = 2$) would be weak evidence against fairness 
but an s-value of 10 would be more persuasive, like getting 10 heads from a set 
of 10 coin flips ($p = 0.5^{10} = 0.0009765625, s = 10$).

Other ideas that look viable to use are reporting minimum false positive risk
(mFPR; Colquhoun, 2009), Bayes Factor Bounds (BFB; Benjamin & Berger, 2019), and 
the upper bound for the posterior probability that that alternate hypothesis
(H1) is true (Benjamin & Berger, 2019). We have opted to use the latter two 
instead of mFPR. 

Taking the Can-Do self-assessment yields an ordinal score ranging from 1-5 that
represents the proficiency level of the learner who took the test. Higher levels
indicate greater proficiency. We can conceptualize the process that yields that
score as a sequential selection process comprised of a set of four level
transition testlets (LTTs) that must be passed in strict order. Every learner
starts at level 1 and only advances to level 2 by passing the first LTT. The
learner's proficiency level is incremented for each LTT passed. The
self-assessment ends as soon as the learner fails to pass an LTT, or reaches the
maximum proficiency level (level 5). For a sample of $N$ learners, there will be
between $N$ and $4N$ binary LTT results (0 = fail, 1 = pass) depending on how 
many learners passed each LTT. 

We use continuation-ratio models to examine the proficiency levels achieved by 
the learners. These are simply logistic regression models applied to a 
reorganized dataset with one row of data per LTT attempted by each learner. 
We are using a predictive modeling approach here that tests how well OPIc scores
and course predict self-assessed proficiency levels. Therefore, we evaluate the
models according to both discrimination and calibration criteria relevant to 
predictive models (Fenlon, O'Grady, Doherty, & Dunnion, 2018). 

\FloatBarrier

## Remaining Tasks

* Consider adding deviance plots based on Fenlon et al. (2018). 
* Streamline flow and eliminate unneeded output. 
* Recheck table object names and resulting table numbering in knitted output. 
* Review and update interpretation text and conclusions. 
* Add code to export figures that will be submitted with the manuscript. 

\FloatBarrier

# Setup
Set global R chunk options (local chunk options will over-ride global options). 

``` {r global-options, cfsize = "footnotesize"}
# Create a custom chunk hook/option for controlling font size in chunk & output.
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$cfsize != "normalsize", paste0("\n \\", options$cfsize,"\n\n", 
                                              x, "\n\n \\normalsize"), x)
  })

# Global chunk options (over-ridden by local chunk options)
knitr::opts_chunk$set(include  = TRUE, echo = TRUE, error = TRUE, 
                      message = TRUE, warning = TRUE, fig.pos = "!ht", 
                      cfsize = "footnotesize")

# Declare location of this script relative to the project root directory.
here::i_am(path = "inst/SAW_Paper_Analyze_Data.Rmd")
```

\FloatBarrier

# LOAD PACKAGES & SET PACKAGE OPTIONS
Some of the functions we use below come from a personal R package called 
[*piercer*](https://github.com/sjpierce/piercer) that is available online from 
[GitHub](https://github.com). Please read and follow the installation 
instructions for *piercer* at https://github.com/sjpierce/piercer
before knitting this script, otherwise the next chunk will not load that 
package and various parts of this script will not work.  

Load contributed R packages that we need to get additional functions. 

``` {r load-packages}
library(here)              # for here()
library(polycor)           # for hetcor()
library(car)               # for residualPlots(), influenceIndexPlot(), outlierTest()
library(multcomp)          # for glht()
library(visreg)            # for visreg()
# Set package options. 
# options(knitr.kable.NA = '0.00')
library(rmarkdown)         # for render(), pandoc_version(). 
library(knitr)             # for kable()
library(texreg)            # for texreg()
library(pander)            # for pander()
library(pROC)              # for roc()
library(ggplot2)           # for ggplot()
library(tidyr)             # for unite()
library(dplyr)             # for filter(), select(), etc.
library(directlabels)      # for direct.label()
library(lattice)           # for strip.custom()
library(modEvA)            # for HLfit()
# Set package options. 
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)        # for kable_styling(), add_header_above(), 
                           # column_spec(), collapse_rows(), and landscape()
library(Hmisc)             # for rcorr()
library(piercer)           # for p2s(), p2bfb(), p2pp(), convertp(), lrcm(), 
                           # brier(), r.p(), r.pc(), r.pc(), ci.rpc(), ci.rp(),
                           # git_report(), which_latex()
library(stringr)           # for word
library(SAWpaper)          # for package version number via session_info()
```

\FloatBarrier

# Load Data

The file `SAW_Paper_Data.RData` contains two data frames: `SData` and `TAData`.
The former has the cleaned student-level data for Spanish students that is 
suitable for use in correlational analyses. The latter data frame contains data
about the same set of students but has been expanded to one row per student per
self-assessment testlet attempted. This is suitable for use in 
continuation-ratio models. 

```{r load-data}
load(file = here::here("data/SAW_Paper_Data.RData"))
```

There are `r nrow(SData)` rows and `r ncol(SData)` variables in `SData`, but 
`r nrow(TAData)` rows and `r ncol(TAData)` variables in `TAData`. 

\FloatBarrier

# DATA MANAGEMENT & EXPLORATION
In `SData`, `Level` (and `LevelF`, which is just a copy of Level that is
stored as a factor instead of a numeric variable) denote the five levels of
Can-Do statements on the self-assessment, whereas in `TAData` the `Testlet`
variable represents the transition testlets that control passage from one level
to the next. The relationship between these variables is illustrated in this
diagram: 1 --> 2 --> 3 --> 4 --> 5. Level in `SData` is represented by the
numbers, and Testlet in `TAData` is represented by the arrows connecting the
numbers. In total, there are 4 transitions (Level in `TAData`) among the 5 levels
(Level in `SData`) of can-do statements on the self-assessment.

Unlike `SData` where there is one row per learner with Level indicating the
maximum level a learner reached on the self-assessment, `TAData` has been expanded
such that there is one row per learner per transition testlet attempted. The binary
variable, Pass, in `TAData` indexes whether a learner passed/failed to pass a
given transition testlet. For instance, a learner who reached Level 3 on the
self-assessment would have one row in `SData` with Level equal to 3. The same 
learner would have three rows in `TAData`, having a 1 on Pass for the rows 
representing testlets 1 and 2, and a 0 on Pass = 0 for the row
representing testlet 3, which was highest transition attempted.

The data in `TAData` is set up as described above for the application of
continuation-ratio modeling, which models the probabilities of passing the four
transitions (represented by Testlet in `TAData`) as a function of predictors (which,
in the case of our study, are course and OPIc speaking scores). Each of those
transitions will eventually yield an estimate of the conditional pass rate,
which is the probability that an individual who reached the level on the left
end of the arrow succeeds in passing on to the level on the right end.
Continuation-ratio models can be conducted as logistic regressions where the
data frame has one row of data per person for each level transition that the
individual actually attempted.

We treat OPIc speaking test proficiency scores as a continuous covariate that
ranges from Novice-low to Superior. The variable OPIc in df_corr_full shows the
actual OPIc ratings. Numerically transfomed OPIc scores are captured by OPIC in
df_corr_full (NL = 1, NM = 2, NH = 3, ..., AM = 8, AH = 9, Superior = 10). We removed
the cases for which the OPIc ratings were not meaningful for this study (AR, BR,
and UR). To make the model coefficients more interpretable, we included centered
OPIc scores instead of the original OPIc results in df_crm_full and stored them as
COPIC. A score of zero in COPIC represents the level of Intermediate-mid on the
OPIc scale.

As readers may have noticed, some students with OPIc scores did not have any 
numerically transformed OPIc values (i.e., NA). This is because these students 
did not receive a valid score on the test. Specifically, these students received 
either above range (AR), below range (BR), or unratable (UR) on the OPIc. AR 
was most likely awarded when a student selected an test form that was well below 
their oral proficiency level. BR was most likely awarded when a student selected 
a test form that was well above their proficiency level. And UR was given when a
student's oral response was not ratable due to one of a variety of reasons 
(e.g., no response, technology failure). For transparency purposes, we include
those students with non-scored OPIc tests in the starting datasets, but we have 
to exclude them from the main analyses because their proficiency levels as 
measured by the OPIc were unknown. In total, 64 students were excluded (4 AR, 
59 BR, and 1 UR). The final analytic sample was composed of 807 students. 

```{r}
# Obtain descriptive statistics for students with non-scored OPIc tests
df_corr_full %>%
  count (OPIc) %>%
  mutate (percentage = paste0(round(n/871, 4)*100, '%'))

# Look at the self-assessment levels of students who got AR, BR, or UR
df_corr_full %>%
  filter (OPIc %in% c("AR", "BR","UR")) %>%
  group_by(OPIc) %>%
  count(Level)
```

Next, we obtain the final analytic samples by excluding students with invalid
OPIc scores.

```{r}
# Final analytic dataset for obtaining descriptive statistics and correaltions
df_corr <- df_corr_full %>% filter (!OPIc %in% c("AR","BR","UR"))
# Final analytic dataset for conducting continuation-ratio modeling
df_crm <- df_crm_full %>% filter (!OPIc %in% c("AR","BR","UR"))
```


\FloatBarrier

We treat the variables of Course, Level, and Testlet in as factor variables that 
are effectively ordinal variables.

We treat OPIc speaking test proficiency scores as a continuous covariate that
ranges from Novice-low to Superior ((NL = 1, NM = 2, NH = 3, ..., AM = 8, 
AH = 9, Superior = 10). The variable `OPICN` in `SData` shows the OPIc ratings
in numeric form and has value labels attached that show the level names. We
already removed the cases for which the OPIc ratings were not meaningful for
this study (AR, BR, and UR) before archiving the file. To make model
coefficients more interpretable, we included centered OPIc scores instead of the
original OPIc results in `TAData` and stored them as `COPIC`. A score of zero in
COPIC represents the Intermediate-mid level on the OPIc scale.

\FloatBarrier

# DATA VISUALIZATION
We visualize the relationship of SA level with OPIc scores and course level 
using scatterplots. We put students' OPIc scores on the x-axis and their total
SA scores (ranging from 1 to 200) on the Y-axis. We assigned different colors to
individual data points based on course level. We also jittered the data 
points so that overlapping points are nudged apart and can be seen more clearly. 

Additionally, we imposed lines onto the scatterplots to better illustrate the 
shape of the relationship between SA level and OPI scores.

In Figure \ref{fig:plot-SAScore-OPICN}, we impose a loess line to represent the
overall relationship between SA level and OPIc scores. A loess line is a smooth
line based on local regression of a dependent variable (Y-variable) on an
independent variable (X- variable). It helps one to see the true relationship
between two variables. In Figure \ref{fig:plot-SAScore-OPICN-by-Course}, we
impose four loess lines to represent the relationship between SA level and OPIc
scores within each of the four course levels. In Figure
\ref{fig:plot-SAScore-OPICN-linear}, we impose a linear regression line to
visually check to what extent the linear line represents the overall
relationship between SA level and OPIc scores. In Figure
\ref{fig:plot-SAScore-OPICN-by-Course-linear}, we impose four linear regression
lines to visually check to what extent the linear lines represent the
relationship between SA level and OPIc scores within each of the four course
levels.

```{r plot-SAScore-OPICN}
# Smooth loess lines added to jittered and colored-coded scatterplots
ggplot(SData, aes(OPICN, Item1_50))+
  geom_jitter(aes(color=Course))+
  geom_smooth(method="loess", se=FALSE, span = .77)+
  xlim(1,9)+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9),
                     labels=c("NL","NM","NH","IL","IM","IH","AL","AM","AH"))+
  scale_y_continuous(breaks=c(40,80,120,160,200),
                     labels=c("Level 1","Level 2","Level 3","Level 4","Level 5"))+
  xlab("OPIc")+
  ylab("Self-Assessment total score")+
  guides(color = guide_legend(title="Course level"))
```

```{r plot-SAScore-OPICN-by-Course}
ggplot(SData, aes(OPICN, Item1_50, color=Course))+
  geom_jitter()+
  geom_smooth(method="loess", se=FALSE, span = .93)+
  xlim(1,9)+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9),
                     labels=c("NL","NM","NH","IL","IM","IH","AL","AM","AH"))+
  scale_y_continuous(breaks=c(40,80,120,160,200),
                     labels=c("Level 1","Level 2","Level 3","Level 4","Level 5"))+
  xlab("OPIc")+
  ylab("Self-assessment can do statements")+
  guides(color = guide_legend(title="Course level"))
```

```{r plot-SAScore-OPICN-linear}
# Linear regression lines added to jittered and colored-coded plots
ggplot(SData, aes(OPICN, Item1_50))+
  geom_jitter(aes(color=Course))+
  geom_smooth(method="lm", se=FALSE)+
  xlim(1,9)+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9),
                     labels=c("NL","NM","NH","IL","IM","IH","AL","AM","AH"))+
  scale_y_continuous(breaks=c(40,80,120,160,200),
                     labels=c("Level 1","Level 2","Level 3","Level 4","Level 5"))+
  xlab("OPIc")+
  ylab("Self-assessment can do statements")+
  guides(color = guide_legend(title="Course level"))
```

```{r plot-SAScore-OPICN-by-Course-linear}
ggplot(SData, aes(OPICN, Item1_50, color=Course))+
  geom_jitter()+
  geom_smooth(method="lm", se=FALSE)+
  xlim(1,9)+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9),
                     labels=c("NL","NM","NH","IL","IM","IH","AL","AM","AH"))+
  scale_y_continuous(breaks=c(40,80,120,160,200),
                     labels=c("Level 1","Level 2","Level 3","Level 4","Level 5"))+
  xlab("OPIc")+
  ylab("Self-assessment can do statements")+
  guides(color = guide_legend(title="Course level"))
```

\FloatBarrier

# CORRELATIONS
We examine the correlations among course level, OPIc scores, and SA level using
the dataset `SData` (not in a long data format). As we mentioned earlier, Level
represents the highest level that a student reached on the SA. Polyserial
correlation is calculated for OPIC and Course and for OPIC and Level, whereas
polychoric correlation is calculated for the two ordinal variables, Course and
Level. We included both ordinal and continuous versions of selected variables to
simplify getting both polychoric and polyserial correlations and (for comparison
purposes only) Pearson correlations that are less appropriate for measuring some
relationships. The names of continuous versions of variables that should be
treated as ordinal have ".n" suffixes. We also estimated selected Spearman 
correlations for a similar comparative purpose. 

```{r}
# Create numeric versions of Level and Course.
SData <- SData %>% 
  mutate(OPIC.n = as.numeric(OPICN),
         Level.n = as.numeric(Level),
         Course.n = as.numeric(Course))

# cvars = continuous variables
cvars <- c("OPIC.n", "Item1_50", "Level.n", "Course.n")

# ovars = ordinal variables
ovars <- c("LevelF", "Course")

# Use the hetcor (heterogenous correlations) in "polycor" package to get a      
# matrix with Pearson correlations (if both variables are continuous), 
# polyserial correlations (if one is continuous and the other is ordinal), and 
# polychoric correlations (if both are ordinal). Warning messages from hetcor() 
# occur because polyserial correlations between ordinal and continuous versions
# of the same variable are not sensible. It was just faster to estimate them as
# part of a single larger matrix than individually get the subset we want. 

HC <- hetcor(as.data.frame(SData[, c(cvars, ovars)]), ML=TRUE, 
             use="pairwise.complete.obs")
rbind(r.ps(HC, cont = cvars, ord = ovars, digits = 2, pdigits = NULL),
      r.pc(HC, ord = ovars, digits = 2, pdigits = NULL)) %>% 
  mutate(Variables = word(rownames(.), 2, sep = "\\:"),
         Type = word(rownames(.), 1, sep = "\\:"),
         Pval = format(Pval, digits = 3)) %>% 
  select(Variables, Type, Cor, SE, CI.LL, CI.UL, Z, Pval, Sval, BFB, PPH1) -> 
  T07

kable(T07, format = "latex", booktabs = TRUE, 
      digits = c(Inf, Inf, rep(x = 2, times = 5), Inf, c(2, 2, 2)),
      caption = "Polyserial and Polychoric Correlations") %>%
kable_styling(latex_options = c("repeat_header"))
```

Obtain Pearson correlations among two continuous variables, OPIC and Item1_50 
(SA sum scores ranging from 1 to 200), and two ordinal variables, Course and 
Level. Note that it is inappropriate to use Pearson's r for ordinal variables. 
We include them here only for comparison purposes. 

```{r pearson}
T08a <- r.p(HC, cont = cvars, digits = 2, pdigits = NULL)
```

Now obtain Spearman correlations for comparison purposes even though they are 
not technically appropriate for our purpose. 

```{r spearman}
T08b <- ci.rp(r = cor(x = SData$Level.n, y = SData$Course.n, 
                      method = "spearman"), 
              n = nrow(SData), rn = "r.s: Level and Course")
T08c <- ci.rp(r = cor(x = SData$Level.n, y = SData$OPIC.n, 
                      method = "spearman"), 
              n = nrow(SData), rn = "r.s: Level and OPIC")
T08d <- ci.rp(r = cor(x = SData$Course.n, y = SData$OPIC.n, 
                      method = "spearman"), 
              n = nrow(SData), rn = "r.s: Course and OPIC")
```

```{r show-correlations}
# Bind into a data frame 
rbind(T08a, T08b, T08c, T08d) %>% 
  mutate(Variables = word(rownames(.), 2, sep = "\\:"),
         Type = word(rownames(.), 1, sep = "\\:"),
         Pval = format(Pval, digits = 3),
         BFB = format(BFB, digits = 3))  %>% 
  select(Variables, Type, Cor, SE, CI.LL, CI.UL, t, df, Pval, Sval, BFB, 
         PPH1) -> 
  T08

kable(T08, format = "latex", booktabs = TRUE, 
      digits = c(rep(x = 2, times = 6), Inf, c(2, 2, 2)),
      caption = "Pearson and Spearman Correlations") %>%
kable_styling(latex_options = c("repeat_header"))
```

Notice that the t and z-statistics are all really large, so the p-values are
zero. That causes all the s-values to become infinite, and the BFB values to
become NaN (not-a-number) because we are dividing by zero in the denominator.
You can read more about interpreting s-values and BFB values in Greenland
(2019), Benjamin & Berger (2019), and Wasserstein, Schirm, & Lazar (2019).

\FloatBarrier

# FIT CONTINUATION-RATIO MODELS
Here we fit a series of continuation-ratio models and display a combined table
summarizing them for comparison purposes. In subsequent sections, we examine 
each of the fitted models in detail. 

```{r fit-models}
# Model 1a: Parallel course effect
m1a <- glm(pass ~ Level + Course - 1, data = TAData, family = "binomial")

# Model 1b: Non-parallel course effect
m1b <- glm(pass ~ Level + Course + Level*Course - 1, data = TAData, 
           family = "binomial")

# Model 2a: Parallel OPIC effect
m2a <- glm(pass ~ Level + COPIC - 1, data = TAData, family = "binomial")

# Model 2b: Non-parallel OPIC effect
m2b <- glm(pass ~ Level + COPIC + Level*COPIC - 1, data = TAData, 
           family = "binomial")

# Model 3a: Parallel OPIC + parallel course effect
m3a <- glm(pass ~ Level + COPIC + Course - 1, data = TAData, 
           family = "binomial")

# Model 3b: Non-parallel OPIC + parallel course effect
m3b <- glm(pass ~ Level + COPIC + Level:COPIC + Course - 1, data = TAData, 
           family = "binomial")

# Create a list of the model fit objects.
TModels <- list(m1a, m1b, m2a, m2b, m3a, m3b)
```

The table below shows the model parameters for the whole set of models.  

\begin{landscape}

```{r display-models, results='asis'} 

texreg(TModels, use.packages = FALSE, dcolumn = TRUE, booktabs = TRUE, 
       single.row = TRUE, digits = 3, label = "TModels", 
       custom.model.names = c("Model 1a",  "Model 1b", "Model 2a", "Model 2b",
                              "Model 3a", "Model3b"), 
       caption = "Continuation-Ratio Models")
```

\end{landscape}

\FloatBarrier

# MODEL 1A: PARALLEL COURSE EFFECT
We fit a series of continuation-ratio models to `TAData` to see whether course
level and OPIc scores significantly predict the pass rate for each transition on
the SA. We first fit a basic model that omits the intercept term in order to
simplify post-processing of the model results into interpretable estimates. We
focus here on a model examining whether the course a learner is taking affects
the pass rate for each level transition. This model assumes that the higher
level courses have a constant effect on the pass rates across all level
transitions. The summary below shows the actual model fit information and the
raw parameter estimates.

\newpage

```{r}
T10 <- summary(m1a)
panderOptions('keep.trailing.zeros', TRUE)
set.caption("Model 1a Summary")
pander(T10, digits = c(5, 5, 4, 3), round = c(4, 4, 2, Inf), justify = "right",
       caption = "Model 1a Summary")
```

Post-process the raw model output to get additional information, including 
s-values, BFBs, and posterior probabilities of H1 corresponding to the p-values 
listed above.

```{r}
format(convertp(T10$coefficients[,"Pr(>|z|)"], digits = 2), digits = 3)
```

\FloatBarrier

## Sequential Tests (Type I SS)
Each row in Table 12 tests the significance of unique additional variance
explained by the term on that line after controlling for all previously entered
terms. Significant results mean adding that term improved the model.

\newpage

```{r}
T12 <- anova(m1a, test = "Chisq")
T12 <- cbind(T12, convertp(T12[,"Pr(>Chi)"]))
pander(T12, justify = "right", split.tables = Inf, 
       digits = c(5, 5, 5, 8, 3, 4, 3, 3), 
       round = c(0, 3, 0, 3, Inf, 2, 2, 2), 
       caption = "Model 1a Sequential Tests (Type I SS): Analysis of Deviance")
```

\FloatBarrier

## Simultaneous Tests of Main Effects via LRT (Type III SS)
The simultaneous tests in Table 13 are the effects of the indicated terms after
controlling for all other terms in the model. They are only computed for terms
that are not part of a higher-order interaction because it makes no sense to
test for a main effect when the variable is involved in an interaction. These
should be functionally equivalent to the results you get with *anova()* when you
feed it a pair of nested models that differ only in that one model includes a
term that is absent from the other model. They are likelihood ratio tests
(LRTs).

```{r}
T13 <- drop1(m1a, test = "Chisq")   
T13 <- cbind(T13, convertp(T13[,"Pr(>Chi)"]))
pander(T13, justify = "right", split.tables = Inf, 
       digits = c(5, 8, 8, 8, 3, 4, 3, 3),
       round = c(0, 3, 3, 3, Inf, 2, 2, 2), 
       caption = "Model 1a Simultaneous Tests (Type III SS)")
```

\FloatBarrier

## Conditional and Unconditional Pass Rates
We use the inverse logit transformation to convert fitted values and associated
confidence intervals into the conditional probability of passing a particular
level transition given that the student is in a particular Spanish course.

We compute the unconditional pass rates from the conditional pass rates as sets
of cumulative products.

```{r}
# Create a new data frame object for use with predict()
m1a.ND <- data.frame(Level = gl(n = 4, k = 1, length = 16,
                               labels = c("1", "2", "3", "4")), 
                    Course = gl(n = 4, k = 4, length = 16, 
                                labels = c("100", "200", "300", "400")))

# Compute predicted mean passing rate at each combination of Level & Course
m1a.pred <- predict(m1a, newdata = m1a.ND, type = "link", se.fit = TRUE)

# Add fitted values and CIs to the new data frame & display it.
critval       <- qnorm(0.975)  # For Wald 95% CIs
m1a.ND$fit    <- m1a.pred$fit
m1a.ND$se.fit <- m1a.pred$se.fit
m1a.ND$fit.LL <- with(m1a.ND, fit - (critval * se.fit))
m1a.ND$fit.UL <- with(m1a.ND, fit + (critval * se.fit))

# Convert fitted values and CIs to probabilities. 
m1a.ND$Pass.Rate <- invlogit(m1a.ND$fit)
m1a.ND$Pass.LL   <- invlogit(m1a.ND$fit.LL)
m1a.ND$Pass.UL   <- invlogit(m1a.ND$fit.UL)

# Compute unconditional pass rates.
m1a.ND$Pass.URate <- c(cumprod(m1a.ND[m1a.ND$Course == "100", "Pass.Rate"]),
                       cumprod(m1a.ND[m1a.ND$Course == "200", "Pass.Rate"]),
                       cumprod(m1a.ND[m1a.ND$Course == "300", "Pass.Rate"]),
                       cumprod(m1a.ND[m1a.ND$Course == "400", "Pass.Rate"]))

ShowVars <- c("Level", "Course", "Pass.Rate", "Pass.LL", "Pass.UL", "Pass.URate")
T14 <- m1a.ND[, ShowVars] 
kable(T14, format = "latex", booktabs = TRUE, digits = 2, 
      format.args = list(nsmall = 2),
      caption = paste("Conditional Pass Rates with 95 percent CIs and",
                      "Unconditional Pass Rates by Level Transition and Course")) %>%
kable_styling(latex_options = c("repeat_header"))
```

\FloatBarrier

## Odds-Ratios for Course Effect
We exponentiate the course parameters to obtain the odds-ratios showing the
effect of being in a second-, third-, or fourth-year Spanish course instead of a
first-year course. This quantifies the effect of course, which this model
assumes is equal across level transitions. We will test whether that assumption
is reasonable in another model later. We are using contrasts to do pairwise
comparisons and get simultaneous 95% CIs that are both adjusted for multiple
testing via Westfall's (1997) method.

```{r}
# Run multiple comparisons examine the course effect & get adjusted 95% CIs. 
m1a.ct <- glht(m1a, linfct = mcp(Course = "Tukey"))
m1a.mc <- summary(m1a.ct, test = adjusted("Westfall"))
m1a.ci <- confint(m1a.ct, calpha = adjusted_calpha(test = "Westfall"))
T15 <- data.frame(Est   = m1a.mc$test$coefficients, 
                  SE    = m1a.mc$test$sigma, 
                  CI.LL = m1a.ci$confint[, "lwr"],
                  CI.UL = m1a.ci$confint[, "upr"],
                  OR    = exp(m1a.mc$test$coefficients),
                  OR.LL = exp(m1a.ci$confint[, "lwr"]),
                  OR.UL = exp(m1a.ci$confint[, "upr"]),
                  z     = m1a.mc$test$tstat,
                  p     = m1a.mc$test$pvalues,
                  Sval  = p2s(m1a.mc$test$pvalues),
                  BFB   = p2bfb(m1a.mc$test$pvalues),
                  PPH1  = p2pp(m1a.mc$test$pvalues))
kable(T15, format = "latex", booktabs = TRUE, format.args = list(digits = 3),
      digits = c(rep(x = 2, times = 8), Inf, 2, 2, 2), 
      caption = paste("Multiple Comparisons for Course Effect, Westfall (1997)",
      "Adjustment for Multiplicity")) %>%
kable_styling(latex_options = c("repeat_header"))
```

\FloatBarrier

## Assessing Goodness of Fit, Discrimination, and Calibration
Since we are running the continuation-ration model as a logistic regression, all
the stadard methods for assessing goodness of fit for logistic regresion models
apply.

\FloatBarrier

### Hosmer-Lemeshow goodness of fit test. 
The null hypothesis for the Hosmer-Lemeshow test (HLT) is that the data fit the model,
so a significant result is actually undesirable because it says that the data
don't fit the model. This is a measure of calibration (Fenlon et al., 2018), 
which can also be examined with a calibration plot like the one shown below.

The calibration plot uses the same set of bins used for the HLT. We used a
binning method that creates bins by grouping the predicted probabilities into
bins with a minimum width of 0.10 because that way all the values grouped
together are quite similar. The confidence interval for each bin is labeled with
the sample size for the bin. Ideally, the point estimate for each bin is close
to the dashed line, but at a minimum we want all the confidence intervals to
overlap that dashed line.

```{r}
HLT.m1a <- HLfit(m1a, bin.method = "prob.bins", min.prob.interval = 0.1, 
                 main = "Calibration Plot", 
                 xlab = "Predicted Probability (minimum bin interval = 0.10)",
                 ylab = "Observed Probability")
HLT.m1a
```

```{r}
# Get the s-value, BFB, and posterior probability of H1. 
convertp(p = HLT.m1a$p.value, digits = 2)
```

\FloatBarrier

### Brier Score
The Brier score is an overall measure of accuracy suitable for logistic 
regression models with binary outcomes (Fenlon et al., 2018). It is the average
prediction error (Steyerberg et al., 2001; Steyerberg et al., 2010). The scaled 
Brier score adjusts for unscaled version to have a range of from 0 to 1, thereby
making it similar to an $R^{2}$ statistic. Values close to 1 are desirable and 
indicate good calibration. We report the scaled Brier score below.

```{r}
brier(m1a, digits = 2)
```

\FloatBarrier

### Classification 
Below, we use functions from the pROC package to identify the optimal cutpoint 
$c$ to use when converting fitted probabilities of passing a LTT into a binary
prediction of whether the individual will pass it. The naieve cutpoint would be
$c = 0.5$, but that is not necessarily the optimal cutpoint. Choosing a cutpoint
via use of the Youden index (Youden, 1950) is optimal with respect to overall
misclassification rate for a given weighting of sensitivity and specificity
(Perkins & Schisterman, 2006). Below, we show the results of various 
classification measures given that we use the optimal cutpoint based on the 
Youden index. 

```{r}
TAData$pred.m1a <- predict(m1a, type = "response")
set.seed(4921) # For reproducibility of bootstrap estimates.
roc.m1a <- roc(pass ~ pred.m1a, data = TAData, ci = TRUE, direction = "<",
               ci.method = "bootstrap", boot.n = 10000)
round(lrcm(roc.m1a, seed = 4684), digits = 3)
```

Here are short definitions of what some of these statistics measure. 

* Accuracy is the model's overall ability to correctly classify learners
  according to whether they pass vs fail the level transitions. 
* Sensitivity is the ability of the model to correctly identify learners who 
  passed level transitions. 
* Specificity is the ability of the model to correctly identify learners who 
  fail the level transitions. 

\FloatBarrier

### Area Under the Curve (AUC) 
The area under the receiver operating characteristic curve (AUC) measures the
model's ability to discriminate those who pass from those who fail to pass the
level transitions. It can range from 0.50 to 1.00; values of 0.50--0.69 are
poor, 0.70-0.79 are acceptable, .80-.89 are excellent, and $\geq$ 0.90 are
outstanding. These interpretive heuristics come from Hosmer, Lemeshow, &
Sturdivant (2013, p. 177).

```{r}
print(roc.m1a)
plot.roc(roc.m1a, print.auc = TRUE, main = "ROC Curve, Model 1a", 
         print.thres = "best")
```

Based on this AUC result, we conclude that model 1a is generally poor at
discriminating people who are passing vs failing the level transitions. That
means course is not a particularly good predictor regardless of what hypothesis
testing results show.

\FloatBarrier

### $R^{2}$ Measures
Pseudo-$R^{2}$ ($R^{2}_p$) is the squared Pearson correlation between observed
and predicted values, as suggested in Hosmer, Lemeshow, & Sturdivant (2013, p.
182). Meanwhile, $R^{2}_{Dev}$ is a measure based on deviance residuals (Fox,
1997, p. 451; Cameron & Windmeijer, 1997).

```{r}
c(PseudoR2 = pseudoR2(m1a, digits = 2), R2Dev = R2Dev(m1a, digits = 2))
```

The low $R^{2}$ is not very encouraging about this model. 

\FloatBarrier

## Diagnostics   
We need to run model diagnostics to see if there are any obvious problems with
the model. First we check for outliers and find that there are none of note.

```{r}
outlierTest(m1a)
```

Now we look at some residual plots.
```{r}
residualPlots(m1a, layout = c(1,3))
```

Now we look at some index plots for influence measures. We are looking for
observations with values exceeding the cutoffs.

```{r}
# Plot Cook's Distance vs observation index.
PlotCookD(m1a)

# How many cases have high Cook's Distances (D > cutoff)?
summary(cooks.distance(m1a))
addmargins(table(cooks.distance(m1a) > CookDco(m1a)))

# Plot Leverage (hat) values vs. observation index.
PlotHat(m1a)

# Plot Standardized Pearson Residuals vs Leverage, w/ Cook's D contours.
plot(m1a, which = 5, id.n = 10, cex.id = .6, 
     cook.levels = round(CookDco(m1a), digits = 3))
abline(v = hatco(m1a), lty = 2, col = "blue")
text(x = hatco(m1a), y = 3.3, pos = 4, col = "blue", cex = .75,
     labels = paste("Leverage cutoff >", round(hatco(m1a), digits = 3)))

# How many cases have high leverage (hat > cutoff)?
summary(hatvalues(m1a))
addmargins(table(hatvalues(m1a) > hatco(m1a)))

# Identify cases w/ high leverage and high Cook's D. 
InfCases(m1a)
```

There are a lot of observations with high Cook's distance and high leverage.
That is a bit concerning, but this may not be the final model we want to use for
drawing conclusions. 

\FloatBarrier

## Graphs

\FloatBarrier

### Main Effects

```{r}
visreg(m1a, ylab = "Log odds (Pass)")
visreg(m1a, ylab = "Conditional Pass Rate", scale = "response", rug = 2)
```

\FloatBarrier

### Conditional Pass Rates
Use ggplot to generate the graph for conditional pass rates derived from model
m1a. 

```{r}
F1 <- ggplot(T14, aes(Level, Pass.Rate, group=Course, color=Course))+
      geom_line()+ coord_cartesian(ylim = c(0, 0.8))+
      xlab("SA Level Transition")+
      ylab("Conditional Pass Rate")+
      ggtitle(label="Conditional Pass Rates by Course and SA Level Transition")
direct.label(F1, "first.polygons")
```

\FloatBarrier

### Unconditional Pass Rates
Use ggplot to generate the graph for unconditional pass rates derived from model
m1a. 

```{r}
F2 <- ggplot(T14, aes(Level, Pass.URate, group=Course, color=Course))+
      geom_line()+ coord_cartesian(ylim = c(0, 0.65))+
      xlab("SA Level Transition")+
      ylab("Unconditional Pass Rate")+
      ggtitle(label="Unconditional Pass Rates by Course and SA Level Transition")
direct.label(F2, "first.polygons")
```

\FloatBarrier

## Conclusion
Learners taking second-, third-, and fourth-year courses are more likely to pass
each of the level transitions than learners in the first-year courses. The
increasing trend in the odds-ratios across the more advanced courses indicate
that the more advanced the course, the larger the difference in the passing
rates is relative to first-year courses. This makes intuitive sense: we expect
learners with more training to do better on self-assessments if the instrument
is a valid measure of Spanish proficiency.

The course effects in this model are assumed to be parallel (constant across the
different level transitions), but we can test whether that assumption needs to
be relaxed by running another model and comparing it to this one.

Despite the significant effects, the model doesn't discriminate well between
those who pass vs. fail (it has modest accuracy, low sensitivity, high
specificity, and low AUC). It explains only a small percentage of the variance.
We can probably do better by considering another predictor. 

\FloatBarrier

# MODEL 1B: NON-PARALLEL COURSE EFFECT
We again omit the intercept term in order to simplify post-processing of the
model results into interpretable estimates. We continue to focus on a model
examining whether the course a learner is taking affects the pass rate, but
relax the assumption that it has a constant effect across all level transitions.
This model allows course to have a level-specific effect on the pass rates. The
summary below shows the actual model fit information and the raw parameter
estimates.

```{r}
T16 <- summary(m1b)
set.caption("Model 1b Summary")
pander(T16, digits = c(5, 5, 4, 3), round = c(4, 4, 2, Inf), justify = "right",
       caption = "Model 1b Summary")
```

Post-process the raw model output to get additional information, including 
s-values, BFBs, and posterior probabilities of H1.

```{r}
format(convertp(T16$coefficients[,"Pr(>|z|)"], digits = 2), digits = 3)
```

Let's show models 1a and 1b side by side. 

```{r, results='asis'} 
T18 <- list(m1a, m1b)
texreg(T18, use.packages = FALSE, dcolumn = TRUE, booktabs = TRUE, 
       single.row = TRUE, digits = 3, label = "T18", 
       custom.model.names = c("Model 1a Parallel",  "Model 1b Non-Parallel"), 
       caption = "Parallel vs. Non-Parallel Course Effect")
```

\newpage

\FloatBarrier

## Sequential Tests (Type I SS)
Here, we want to focus on the result for the interaction effect. That tells us
whether allowing course to have a level-specific effect instead of a constant
effect across levels improved the model.

```{r}
T19 <- anova(m1b, test = "Chisq")
T19 <- cbind(T19, convertp(T19[,"Pr(>Chi)"]))
pander(T19, justify = "right", split.tables = Inf, 
       digits = c(5, 5, 5, 8, 3, 4, 3, 3), 
       round = c(0, 3, 0, 3, Inf, 2, 2, 2), 
       caption = "Model 1b Sequential Tests (Type I SS): Analysis of Deviance")
```

\FloatBarrier

## Simultaneous Tests of Interaction Effects via LRT (Type III SS)

Another way to test whether the interaction was significant is to compare models
via a likelihood ratio test (LRT). We can do that by feeding the *anova()*
function two model fit objects. The models must be nested (one contains only a
subset of the parameters in the other model).

\newpage

```{r}
#CONTINUE  HERE
T20 <- anova(m1a, m1b, test = "Chisq")
T20 <- cbind(T20, convertp(T20[,"Pr(>Chi)"]))
pander(T20, justify = "right", split.tables = Inf, 
       digits = c(5, 8, 5, 8, 3, 4, 3, 3), 
       round = c(0, 3, 0, 3, Inf, 2, 2, 2),
       caption = "Model 1a vs 1b: Likelihood Ratio Test for Interaction")
```

Both methods demonstrated for testing the interaction term yield the same
p-value, so we will only use one of them from here on out.

\FloatBarrier

## Conditional and Unconditional Pass Rates
We omit these rates because model 1b is not better than model 1a. 

\FloatBarrier

## Odds-Ratios for Course Effect
We omit these odds-ratios because model 1b is not better than model 1a. 

\FloatBarrier

## Assessing Goodness of Fit, Discrimination, and Calibration
Since we are running the continuation-ration model as a logistic regression, all
the stadard methods for assessing goodness of fit for logistic regresion models
apply.

\FloatBarrier

### Hosmer-Lemeshow goodness of fit test. 
The null hypothesis for the Hosmer-Lemeshow test is that the data fit the model,
so a significant result is actually undesirable because it says that the data
don't fit the model. This is a measure of calibration (Fenlon et al., 2018), 
which can also be examined with a calibration plot like the one shown below.

The calibration plot uses the same set of bins used for the HLT. We used a
binning method that creates bins by grouping the predicted probabilities into
bins with a minimum width of 0.10 because that way all the values grouped
together are quite similar.

```{r}
HLT.m1b <- HLfit(m1b, bin.method = "prob.bins", min.prob.interval = 0.1, 
                 main = "Calibration Plot", 
                 xlab = "Predicted Probability (minimum bin interval = 0.10)",
                 ylab = "Observed Probability")
HLT.m1b
```

```{r}
# Get the s-value, BFB, and posterior probability of H1. 
convertp(p = HLT.m1b$p.value, digits = 2)
```

\FloatBarrier

### Brier Score
The unscaled Brier score is an overall measure of accuracy suitable for logistic
regression models with binary outcomes (Fenlon et al., 2018). It is the average
prediction error (Steyerberg et al., 2001; Steyerberg et al., 2010). The scaled 
Brier score adjusts for unscaled version to have a range of from 0 to 1, thereby
making it similar to an $R^{2}$ statistic. We report the scaled Brier score 
below.  

```{r}
brier(m1b, digits = 2)
```

\FloatBarrier

### Classification 
Here we want to examine the classification table, plus the overall accuracy,
sensitivity, and specificity estimates.

```{r}
TAData$pred.m1b <- predict(m1b, type = "response")
set.seed(1574) # For reproducible bootstrap estimates.
roc.m1b <- roc(pass ~ pred.m1b, data = TAData, ci = TRUE, direction = "<",
               ci.method = "bootstrap", boot.n = 10000)
round(lrcm(roc.m1b, seed = 6711), digits = 3)
```

The low sensitivity indicates that the model does a poor job of correctly
identifying those who pass the level transitions. The high specificity indicates
that it does a good job of identifying those who fail to pass level transitions.

\FloatBarrier

### Area Under the Curve (AUC)
The area under the receiver operating characteristic curve (AUC) measures the
model's ability to discriminate those who pass from those who fail to pass the
level transitions. It can range from 0.50 to 1.00; values of 0.50--0.69 are
poor, 0.70-0.79 are acceptable, .80-.89 are excellent, and $\geq$ 0.90 are
outstanding.

```{r}
print(roc.m1b)
plot.roc(roc.m1b, print.auc = TRUE, main = "ROC Curve, Model 1b", 
         print.thres = "best")
```

Based on this AUC result, we conclude that model 1b does a poor job of
discriminating between those who pass vs fail the level transitions. That means
course is still not a particularly good predictor.

Next we compare the ROC curves for models m1a and m1b using a bootstrap test for
difference between the AUC values (Robin et al., 2011). 

```{r}
set.seed(9537) # For reproducible bootstrap estimates.
roct.m1a.m1b <- roc.test(roc.m1a, roc.m1b, method = "bootstrap", boot.n = 10000)
roct.m1a.m1b
```

```{r}
# Get the s-value, BFB, and posterior probability of H1. 
convertp(p = roct.m1a.m1b$p.value, digits = 2)
```

\FloatBarrier

### $R^{2}$ Measures
Pseudo-$R^{2}$ ($R^{2}_p$) is the squared Pearson correlation between observed
and predicted values, as suggested in Hosmer, Lemeshow, & Sturdivant (2013, p.
182). Meanwhile, $R^{2}_{Dev}$ is a measure based on deviance residuals (Fox,
1997, p. 451; Cameron & Windmeijer, 1997). The result for model 1b is no better
than it was for model 1a.

```{r}
c(PseudoR2 = pseudoR2(m1b, digits = 2), R2Dev = R2Dev(m1b, digits = 2))
```

\FloatBarrier

## Diagnostics
We omit detailed diagnostics because model 1b is not better than model 1a. 

\FloatBarrier

## Graphs
We omit graphs because model 1b is not better than model 1a. See model 1a graphs
instead.

\FloatBarrier

## Conclusion
The interaction is not improving the model. We should interpret model 1a instead
of model 1b. That may change if we build in other covariates. However, we should
also test COPIC as an alternative to course. We turn to that next in model 2a.

\FloatBarrier

# MODEL 2A: PARALLEL OPIC EFFECT
We now want to look at a different covariate, namely OPIc speaking proficiency
scores. We again fit a model that omits the intercept term in order to simplify
post-processing of the model results into interpretable estimates. We start by
assuming that the OPIc scores have a constant effect on the pass rates across
all level transitions. We are using the centered version of OPIc in the model.
The summary below shows the actual model fit information and the raw parameter
estimates.

```{r}
T21 <- summary(m2a)
set.caption("Model 2a Summary")
pander(T21, digits = c(5, 5, 4, 3), round = c(4, 4, 2, Inf), justify = "right",
       caption = "Model 2a Summary")
```

Post-process the raw model output to get additional information, including 
s-values, BFBs, and posterior probabilities of H1.

```{r}
format(convertp(T21$coefficients[,"Pr(>|z|)"], digits = 2), digits = 3)
```

\FloatBarrier

# Sequential Tests (Type I SS)
Each row in Table 23 tests the significance of unique additional variance
explained by the term on that line after controlling for all previously entered
terms. Significant results mean adding that term improved the model.

```{r}
T23 <- anova(m2a, test = "Chisq")
T23 <- cbind(T23, convertp(T23[,"Pr(>Chi)"]))
pander(T23, justify = "right", split.tables = Inf, 
       digits = c(5, 5, 5, 8, 3, 4, 3, 3), 
       round = c(0, 3, 0, 3, Inf, 2, 2, 2),
       caption = "Model 2a Sequential Tests (Type I SS): Analysis of Deviance")
```

\FloatBarrier

## Simultaneous Tests of Main Effects via LRT (Type III SS)
The simultaneous tests in Table 24 are the effects of the indicated terms after
controlling for all other terms in the model. They are only computed for terms
that are not part of a higher-order interaction because it makes no sense to
test for a main effect when the variable is involved in an interaction. These
should be functionally equivalent to the results you get with *anova()* when you
feed it a pair of nested models that differ only in that one model includes a
term that is absent from the other model. They are likelihood ratio tests
(LRTs).

```{r}
T24 <- drop1(m2a, test = "Chisq") 
T24 <- cbind(T24, convertp(T24[,"Pr(>Chi)"]))
pander(T24, justify = "right", split.tables = Inf, 
       digits = c(5, 8, 8, 8, 3, 4, 3, 3),
       round = c(0, 3, 3, 3, Inf, 2, 2, 2), 
       caption = "Model 2a Simultaneous Tests (Type III SS)")
```

\FloatBarrier

## Conditional and Unconditional Pass Rates
We use the inverse logit transformation to convert fitted values and associated 
confidence intervals into the conditional probability of passing a particular 
level transition given that the student has a particular COPIC score.
We compute the unconditional pass rates from the conditional pass rates as sets 
of cumulative products.

```{r}
# Create a new data frame object for use with predict()
m2a.ND <- data.frame(Level = gl(n = 4, k = 1, length = 36, 
                                labels = c("1", "2", "3", "4")),
                     COPIC = rep(-4:4, each = 4), OPIC = rep(1:9, each = 4))

# Compute predicted mean passing rate at each combination of Level & Course
m2a.pred <- predict(m2a, newdata = m2a.ND, type = "link", se.fit = TRUE)

# Add fitted values and CIs to the new data frame & display it.
critval       <- qnorm(0.975) # For Wald 95% CIs
m2a.ND$fit    <- m2a.pred$fit
m2a.ND$se.fit <- m2a.pred$se.fit
m2a.ND$fit.LL <- with(m2a.ND, fit - (critval * se.fit))
m2a.ND$fit.UL <- with(m2a.ND, fit + (critval * se.fit))

# Convert fitted values and CIs to probabilities. 
m2a.ND$Pass.Rate <- invlogit(m2a.ND$fit)
m2a.ND$Pass.LL   <- invlogit(m2a.ND$fit.LL)
m2a.ND$Pass.UL   <- invlogit(m2a.ND$fit.UL)

# Compute unconditional pass rates.
m2a.ND$Pass.URate <- c(cumprod(m2a.ND[m2a.ND$COPIC == -4, "Pass.Rate"]),
                       cumprod(m2a.ND[m2a.ND$COPIC == -3, "Pass.Rate"]),
                       cumprod(m2a.ND[m2a.ND$COPIC == -2, "Pass.Rate"]),
                       cumprod(m2a.ND[m2a.ND$COPIC == -1, "Pass.Rate"]),
                       cumprod(m2a.ND[m2a.ND$COPIC == 0, "Pass.Rate"]),
                       cumprod(m2a.ND[m2a.ND$COPIC == 1, "Pass.Rate"]),
                       cumprod(m2a.ND[m2a.ND$COPIC == 2, "Pass.Rate"]),
                       cumprod(m2a.ND[m2a.ND$COPIC == 3, "Pass.Rate"]),
                       cumprod(m2a.ND[m2a.ND$COPIC == 4, "Pass.Rate"]))

ShowVars <- c("Level", "COPIC", "OPIC", "Pass.Rate", "Pass.LL", "Pass.UL", "Pass.URate")
T25 <- m2a.ND[, ShowVars] 
kable(T25, format = "latex", booktabs = TRUE, digits = 2, 
      format.args = list(nsmall = 2),
      caption = paste("Conditional Pass Rates (with 95 percent CIs) and",
                      "Unconditional Pass Rates by Level Transition and Course")) %>% 
kable_styling(latex_options = c("repeat_header")) 
```

\FloatBarrier

## Odds-Ratio for COPIC Effect 
We exponentiate the COPIC parameter to obtain the odds-ratio showing the effect
of an extra point on the COPIC (i.e., a score of 6 vs 5 in the raw variable).
This quantifies the effect of COPIC, which this model assumes is equal across
level transitions. We will test whether that assumption is reasonable in another
model later.

```{r}
T26   <- cbind(OR = exp(coef(m2a)[5]), 
               LL = exp(confint(m2a, level = 0.95)[5, 1]),
               UL = exp(confint(m2a, level = 0.95)[5, 2]))
dimnames(T26)[[1]] <- "COPIC (0 vs 1)"
kable(T26, format = "latex", booktabs = TRUE, digits = 2, 
      caption = paste("Odds-Ratios for COPIC Effect (Estimates and 95 Percent",
                      "Confidence Intervals)")) %>% 
kable_styling(latex_options = c("repeat_header")) 
```

\FloatBarrier

## Assessing Goodness of Fit, Discrimination, and Calibration
Since we are running the continuation-ration model as a logistic regression, all
the stadard methods for assessing goodness of fit for logistic regresion models
apply.

\FloatBarrier

### Hosmer-Lemeshow goodness of fit test. 
The null hypothesis for the Hosmer-Lemeshow test is that the data fit the model,
so a significant result is actually undesirable because it says that the data
don't fit the model. This is a measure of calibration (Fenlon et al., 2018), 
which can also be examined with a calibration plot like the one shown below.

The calibration plot uses the same set of bins used for the HLT. We used a
binning method that creates bins by grouping the predicted probabilities into
bins with a minimum width of 0.10 because that way all the values grouped
together are quite similar.

```{r}
HLT.m2a <- HLfit(m2a, bin.method = "prob.bins", min.prob.interval = 0.1, 
                 main = "Calibration Plot", 
                 xlab = "Predicted Probability (minimum bin interval = 0.10)",
                 ylab = "Observed Probability")
HLT.m2a
```

```{r}
# Get the s-value, BFB, and posterior probability of H1. 
convertp(p = HLT.m2a$p.value, digits = 2)
```

\FloatBarrier

### Brier Score
The unscaled Brier score is an overall measure of accuracy suitable for logistic
regression models with binary outcomes (Fenlon et al., 2018). It is the average
prediction error (Steyerberg et al., 2001; Steyerberg et al., 2010). The scaled 
Brier score adjusts for unscaled version to have a range of from 0 to 1, thereby
making it similar to an $R^{2}$ statistic. We report the scaled Brier score 
below.  

```{r}
brier(m2a, digits = 2)
```

\FloatBarrier

### Classification 
Here we want to examine the classification table, plus the overall accuracy,
sensitivity, and specificity estimates.

```{r}
TAData$pred.m2a <- predict(m2a, type = "response")
set.seed(5378) # For reproducible bootstrap estimates.
roc.m2a <- roc(pass ~ pred.m2a, data = TAData, ci = TRUE, direction = "<",
               ci.method = "bootstrap", boot.n = 10000)
round(lrcm(roc.m2a, seed = 9825), digits = 3)
```

\FloatBarrier

### Area Under the Curve (AUC)
The area under the receiver operating characteristic curve (AUC) measures the
model's ability to discriminate those who pass from those who fail to pass the
level transitions. It can range from 0.50 to 1.00; values of 0.50--0.69 are
poor, 0.70-0.79 are acceptable, .80-.89 are excellent, and $\geq$ 0.90 are
outstanding.

```{r}
print(roc.m2a)
plot.roc(roc.m2a, print.auc = TRUE, main = "ROC Curve, Model 2a", 
         print.thres = "best")
```

Next we compare the ROC curves for models m1a and m2a using a bootstrap test for
difference between the AUC values (Robin et al., 2011). 

```{r}
set.seed(7436) # For reproducible bootstrap estimates.
roct.m1a.m2a <- roc.test(roc.m1a, roc.m2a, method = "bootstrap", boot.n = 10000)
roct.m1a.m2a
```

```{r}
# Get the s-value, BFB, and posterior probability of H1. 
convertp(p = roct.m1a.m2a$p.value, digits = 2)
```

\FloatBarrier

### $R^{2}$ Measures
Pseudo-$R^{2}$ ($R^{2}_p$) is the squared Pearson correlation between observed
and predicted values, as suggested in Hosmer, Lemeshow, & Sturdivant (2013, p.
182). Meanwhile, $R^{2}_{Dev}$ is a measure based on deviance residuals (Fox,
1997, p. 451; Cameron & Windmeijer, 1997).

```{r}
c(PseudoR2 = pseudoR2(m2a, digits = 2), R2Dev = R2Dev(m2a, digits = 2))
```

\FloatBarrier

## Diagnostics
We need to run model diagnostics to see if there are any obvious problems with 
the model. First we check for outliers and find that there are none of note.

```{r}
outlierTest(m2a)
```

Now we look at some residual plots.
```{r}
residualPlots(m2a, layout = c(1,3))
```

Now we look at some index plots for influence measures. We are looking for
observations with values exceeding the cutoffs.

```{r}
# Plot Cook's Distance vs observation index.
PlotCookD(m2a)

# How many cases have high Cook's Distances (D > cutoff)?
summary(cooks.distance(m2a))
addmargins(table(cooks.distance(m2a) > CookDco(m2a)))

# Plot Leverage (hat) values vs. observation index.
PlotHat(m2a)

# Plot Standardized Pearson Residuals vs Leverage, w/ Cook's D contours.
plot(m2a, which = 5, id.n = 10, cex.id = .6, 
     cook.levels = round(CookDco(m2a), digits = 3))
abline(v = hatco(m2a), lty = 2, col = "blue")
text(x = hatco(m2a), y = 5.7, pos = 4, col = "blue", cex = .75,
     labels = paste("Leverage cutoff >", round(hatco(m2a), digits = 3)))

# How many cases have high leverage (hat > cutoff)?
summary(hatvalues(m2a))
addmargins(table(hatvalues(m2b) > hatco(m2a)))

# Identify cases w/ high leverage and high Cook's D. 
InfCases(m2a)
```

\FloatBarrier

## Graphs

### COPIC Effect in log odds (logits) and conditional pass rates

We visualize below the level-transition-specific OPIc effect. The effect of OPIc
decreases as we go from the lowest to the highest level transition, as evidenced
by the decreasingly steep slope from the left-most to the right-most panel.

```{r}
visreg(m2a, xvar = "COPIC", by = "Level", layout = c(4, 1), jitter = TRUE,
       strip.names = TRUE, scale = "linear", 
       xlab = "Centered OPIC Score (OPIC - 5)", 
       ylab = "Log odds (Pass)")
visreg(m2a, xvar = "COPIC", by = "Level", layout = c(4, 1), jitter = TRUE,
       strip.names = TRUE, scale = "response", rug = 2,
       xlab = "Centered OPIC Score (OPIC - 5)", 
       ylab = "Conditional Pass Rate")
```

\FloatBarrier

### Conditional Pass Rates
Use ggplot to generate the graph for conditional pass rates derived from model
m2a.

```{r}
F3 <- ggplot(T25, aes(Level, Pass.Rate, group=OPIC, color=OPIC))+
  geom_line()+ coord_cartesian(ylim = c(0, 1))+
  xlab("Self-assessment Threshold")+
  ylab("Conditional Pass Rate")+
  ggtitle(label="Conditional Pass Rates by OPIc Score and Self-assessment Threshold")
direct.label(F3, "first.polygons")
# for the manuscript
ggsave("F4.png",plot=direct.label(F3, "first.polygons"), width = 6.5, height = 8.5, dpi = 300)
```

\FloatBarrier

### Unconditional Pass Rates  
Use ggplot to generate the graph for unconditional pass rates derived from model
m2a.

```{r}
F4 <- ggplot(T25, aes(Level, Pass.URate, group=OPIC, color=OPIC))+
  geom_line()+ coord_cartesian(ylim = c(0, 1))+
  xlab("Self-assessment Threshold")+
  ylab("Unconditional Pass Rate")+
  ggtitle(label="Unconditional Pass Rates by OPIc Score and Self-assessment Threshold")
direct.label(F4, "first.polygons")
# for the manuscript
ggsave("F5.png",plot=direct.label(F4, "first.polygons"), width = 9, height = 5.5, dpi = 300)
```

\FloatBarrier

## Conclusion
Model 2a suggests that higher OPIC scores are associated with higher passing
rates across the level transitions. Furthermore, model 2a is clearly better than
model 1a. However, we still need to run another model to see whether relaxing
the assumption that the OPIC effect is constant across levels is viable. We do
that in model 2b.

\newpage

\FloatBarrier

# MODEL 2B: NON-PARALLEL OPIC EFFECT
We now relax the assumption that the OPIC effect is constant across level
transitions. We again fit a model that omits the intercept term in order to
simplify post-processing of the model results into interpretable estimates. We
are still using the centered version of OPIC in the model. The summary below
shows the actual model fit information and the raw parameter estimates.

```{r}
T27 <- summary(m2b)
set.caption("Model 2b Summary")
pander(T27, digits = c(5, 5, 4, 3), round = c(4, 4, 2, Inf), justify = "right",
       caption = "Model 2b Summary")
```

Post-process the raw model output to get additional information, including 
s-values, BFBs, and posterior probabilities of H1.

```{r}
format(convertp(T27$coefficients[,"Pr(>|z|)"], digits = 2), digits = 3)
```

Let's show the models 2a and 2b side by side. 

```{r, results='asis'} 
T29 <- list(m2a, m2b)
texreg(T29, use.packages = FALSE, dcolumn = TRUE, booktabs = TRUE, 
       single.row = TRUE, digits = 3, label = "T26", 
       custom.model.names = c("Model 2a Parallel", "Model 2b Non-Parallel"), 
       caption = "Parallel vs. Non-Parallel OPIC Effect")
```

\newpage

We can see that the difference in AICs
($\Delta AIC = AIC_{2a} - AIC_{2b} = `r round(AIC(m2a) - AIC(m2b), digits = 3)`$)
favors model 2b, which has the lower AIC value, but not by a large margin. 
Meanwhile, difference in BICs 
($\Delta BIC = BIC_{2a} - BIC_{2b} = `r round(BIC(m2a) - BIC(m2b), digits = 3)`$) 
favors model 2a by a larger margin because it imposes a stronger penalty for model complexity 
(lack of parsimony). 

\FloatBarrier

## Sequential Tests (Type I SS)
Each row in Table 30 tests the significance of unique additional variance
explained by the term on that line after controlling for all previously entered
terms. Significant results mean adding that term improved the model. Here, we
want to focus on the result for the interaction effect. That tells us whether
allowing OPIc to have a level-specific effect instead of a constant effect
across levels improved the model.

```{r}
T30 <- anova(m2b, test = "Chisq")
T30 <- cbind(T30, convertp(T30[,"Pr(>Chi)"]))
pander(T30, justify = "right", split.tables = Inf, 
       digits = c(5, 5, 5, 8, 3, 4, 3, 3), 
       round = c(0, 3, 0, 3, Inf, 2, 2, 2), 
       caption = "Model 2b Sequential Tests (Type I SS): Analysis of Deviance")
```

\newpage

\FloatBarrier

## Simultaneous Tests of Interaction Effects via LRT (Type III SS)
Another way to test whether the interaction was significant is to compare models
via a likelihood ratio test (LRT). We can do that by feeding the *anova()*
function two model fit objects. The models must be nested (one contains only a
subset of the parameters in the other model).

```{r}
T31 <- anova(m2a, m2b, test = "Chisq")   
T31 <- cbind(T31, convertp(T31[,"Pr(>Chi)"]))
pander(T31, justify = "right", split.tables = Inf, 
       digits = c(5, 8, 5, 8, 3, 4, 3, 3), 
       round = c(0, 3, 0, 3, Inf, 2, 2, 2),
       caption = "Model 2b Simultaneous Tests (Type III SS)")
```

\FloatBarrier

## Conditional and Unconditional Pass Rates
We use the inverse logit transformation to convert fitted values and associated
confidence intervals into the conditional probability of passing a particular
level transition given that the student has a particular COPIC score.

We compute the unconditional pass rates from the conditional pass rates as sets
of cumulative products.

```{r}
# Create a new data frame object for use with predict()
m2b.ND <- data.frame(Level = gl(n = 4, k = 1, length = 36,
                               labels = c("1", "2", "3", "4")), 
                     COPIC = rep(-4:4, each = 4),
                     OPIC = rep(1:9, each = 4))

# Compute predicted mean passing rate at each combination of Level & Course
m2b.pred <- predict(m2b, newdata = m2b.ND, type = "link", se.fit = TRUE)

# Add fitted values and CIs to the new data frame & display it.
critval      <- qnorm(0.975)  # For Wald 95% CIs
m2b.ND$fit    <- m2b.pred$fit
m2b.ND$se.fit <- m2b.pred$se.fit
m2b.ND$fit.LL <- with(m2b.ND, fit - (critval * se.fit))
m2b.ND$fit.UL <- with(m2b.ND, fit + (critval * se.fit))

# Convert fitted values and CIs to probabilities. 
m2b.ND$Pass.Rate <- invlogit(m2b.ND$fit)
m2b.ND$Pass.LL   <- invlogit(m2b.ND$fit.LL)
m2b.ND$Pass.UL   <- invlogit(m2b.ND$fit.UL)

# Compute unconditional pass rates.
m2b.ND$Pass.URate <- c(cumprod(m2b.ND[m2b.ND$COPIC == -4, "Pass.Rate"]),
                       cumprod(m2b.ND[m2b.ND$COPIC == -3, "Pass.Rate"]),
                       cumprod(m2b.ND[m2b.ND$COPIC == -2, "Pass.Rate"]),
                       cumprod(m2b.ND[m2b.ND$COPIC == -1, "Pass.Rate"]),
                       cumprod(m2b.ND[m2b.ND$COPIC == 0, "Pass.Rate"]),
                       cumprod(m2b.ND[m2b.ND$COPIC == 1, "Pass.Rate"]),
                       cumprod(m2b.ND[m2b.ND$COPIC == 2, "Pass.Rate"]),
                       cumprod(m2b.ND[m2b.ND$COPIC == 3, "Pass.Rate"]),
                       cumprod(m2b.ND[m2b.ND$COPIC == 4, "Pass.Rate"]))

ShowVars <- c("Level", "COPIC", "OPIC", "Pass.Rate", "Pass.LL", "Pass.UL", "Pass.URate")
T32 <- m2b.ND[, ShowVars] 
kable(T32, format = "latex", booktabs = TRUE, digits = 2, 
      format.args = list(nsmall = 2),
      caption = paste("Conditional Pass Rates (with 95 percent CIs) and",
                      "Unconditional Pass Rates by Level Transition and Course")) %>% 
kable_styling(latex_options = c("repeat_header")) 
```

\FloatBarrier

## Odds-Ratios for COPIC Effect    
We exponentiate combinations of the parameters to obtain the odds-ratios showing
the effect of COPIC at each level. This quantifies the effect of COPIC, which
this model assumes varies across level transitions. We are using contrasts that
estimate the simple slope of COPIC at each level, then get simultaneous 95% CIs
that are adjusted for multiple testing via Westfall's (1997) method.

```{r}
# Create objects to hold row names and a contrast matrix (K).
RN <- c("Slope @ L1", "Slope @ L2", "Slope @ L3", "Slope @ L4",
        "Slope diff: L1 - L2", "Slope diff: L1 - L3",
        "Slope diff: L1 - L4", "Slope diff: L2 - L3",
        "Slope diff: L2 - L4", "Slope diff: L3 - L4")
K <- matrix(c(0, 0, 0, 0, 1, 0, 0, 0, 
              0, 0, 0, 0, 1, 1, 0, 0, 
              0, 0, 0, 0, 1, 0, 1, 0, 
              0, 0, 0, 0, 1, 0, 0, 1, 
              0, 0, 0, 0, 0, 1, 0, 0, 
              0, 0, 0, 0, 0, 0, 1, 0, 
              0, 0, 0, 0, 0, 0, 0, 1, 
              0, 0, 0, 0, 0, 1, -1, 0, 
              0, 0, 0, 0, 0, 1, 0, -1, 
              0, 0, 0, 0, 0, 0, 1, -1),
            nrow=10, byrow = TRUE,
            dimnames = list(RN, names(coef(m2b))))

# Run multiple comparisons examine the course effect & get adjusted 95% CIs. 
m2b.ct <- glht(m2b, linfct = K)
m2b.mc <- summary(m2b.ct, test = adjusted("Westfall"))
m2b.ci <- confint(m2b.ct, calpha = adjusted_calpha(test = "Westfall"))
T33 <- data.frame(Est   = m2b.mc$test$coefficients, 
                  SE    = m2b.mc$test$sigma, 
                  CI.LL = m2b.ci$confint[, "lwr"],
                  CI.UL = m2b.ci$confint[, "upr"],
                  OR    = exp(m2b.mc$test$coefficients),
                  OR.LL = exp(m2b.ci$confint[, "lwr"]),
                  OR.UL = exp(m2b.ci$confint[, "upr"]),
                  z     = m2b.mc$test$tstat,
                  p     = m2b.mc$test$pvalues,
                  Sval  = p2s(m2b.mc$test$pvalues),
                  BFB   = p2bfb(m2b.mc$test$pvalues),
                  PPH1  = p2pp(m2b.mc$test$pvalues))
T33$OR[5:10]    <- NA
T33$OR.LL[5:10] <- NA
T33$OR.UL[5:10] <- NA
kable(T33, format = "latex", booktabs = TRUE, format.args = list(digits = 3),
      digits = c(rep(x = 2, times = 8), Inf, 2, 2, 2),
      caption = paste("Simple Slopes of COPIC Effect, Westfall (1997)",
                      "Adjustment for Multiplicity")) %>% 
kable_styling(latex_options = c("repeat_header", "scale_down"))
```

\FloatBarrier

## Assessing Assessing Goodness of Fit, Discrimination, and Calibration
Since we are running the continuation-ration model as a logistic regression, all
the stadard methods for assessing goodness of fit for logistic regresion models
apply.

\FloatBarrier

### Hosmer-Lemeshow goodness of fit test. 
The null hypothesis for the Hosmer-Lemeshow test is that the data fit the model,
so a significant result is actually undesirable because it says that the data
don't fit the model. This is a measure of calibration (Fenlon et al., 2018), 
which can also be examined with a calibration plot like the one shown below.

The calibration plot uses the same set of bins used for the HLT. We used a
binning method that creates bins by grouping the predicted probabilities into
bins with a minimum width of 0.17 because that way all the values grouped
together are quite similar. We tried a bin width of 0.10 as with the other 
models, Unlikebut that yielded bins with fewwer than 15 observations, for which
comparisons may not be meaningful. We aimed for the narrowest bins that had 
adequate size by adding 0.01 to the minimum bin width at each step until all 
bins had $N \ge 15$ observations.

```{r}
HLT.m2b <- HLfit(m2b, bin.method = "prob.bins", min.prob.interval = 0.17, 
                 main = "Calibration Plot", 
                 xlab = "Predicted Probability (minimum bin interval = 0.17)",
                 ylab = "Observed Probability")
HLT.m2b
```


```{r}
# Get the s-value, BFB, and posterior probability of H1. 
convertp(p = HLT.m2b$p.value, digits = 2)
```

\FloatBarrier

### Brier Score
The unscaled Brier score is an overall measure of accuracy suitable for logistic
regression models with binary outcomes (Fenlon et al., 2018). It is the average
prediction error (Steyerberg et al., 2001; Steyerberg et al., 2010). The scaled 
Brier score adjusts for unscaled version to have a range of from 0 to 1, thereby
making it similar to an $R^{2}$ statistic. We report the scaled Brier score 
below.  

```{r}
brier(m2b, digits = 2)
```

\FloatBarrier

### Classification 
Here we want to examine the classification table, plus the overall accuracy,
sensitivity, and specificity estimates.

```{r}
TAData$pred.m2b <- predict(m2b, type = "response")
set.seed(3179) # For reproducible bootstrap estimates.
roc.m2b <- roc(pass ~ pred.m2b, data = TAData, ci = TRUE, direction = "<",
               ci.method = "bootstrap", boot.n = 10000)
round(lrcm(roc.m2b, seed = 7146), digits = 3)
```

\FloatBarrier

### Area Under the Curve (AUC)
The area under the receiver operating characteristic curve (AUC) measures the
model's ability to discriminate those who pass from those who fail to pass the
level transitions. It can range from 0.50 to 1.00; values of 0.50--0.69 are
poor, 0.70-0.79 are acceptable, .80-.89 are excellent, and $\geq$ 0.90 are
outstanding.

```{r}
print(roc.m2b)
plot.roc(roc.m2b, print.auc = TRUE, main = "ROC Curve, Model 2b", 
         print.thres = "best")
```

Next we compare the ROC curves for models m2a and m2b using a bootstrap test for
the difference between the AUC values (Robin et al., 2011).

```{r}
set.seed(1128) # For reproducible bootstrap estimates.
roct.m2a.m2b <- roc.test(roc.m2a, roc.m2b, method = "bootstrap", boot.n = 10000)
roct.m2a.m2b
```

```{r}
# Get the s-value, BFB, and posterior probability of H1. 
convertp(p = roct.m2a.m2b$p.value, digits = 2)
```

\FloatBarrier

### $R^{2}$ Measures
Pseudo-$R^{2}$ ($R^{2}_p$) is the squared Pearson correlation between observed
and predicted values, as suggested in Hosmer, Lemeshow, & Sturdivant (2013, p.
182). Meanwhile, $R^{2}_{Dev}$ is a measure based on deviance residuals (Fox,
1997, p. 451; Cameron & Windmeijer, 1997).

```{r}
c(PseudoR2 = pseudoR2(m2b, digits = 2), R2Dev = R2Dev(m2b, digits = 2))
```

\FloatBarrier

## Diagnostics
We need to run model diagnostics to see if there are any obvious problems with
the model. First we check for outliers and find that there are none of note.

```{r}
outlierTest(m2b)
```

Now we look at some residual plots.
```{r}
residualPlots(m2b, layout = c(1,3))
```

Now we look at some index plots for influence measures. We are looking for
observations with values exceeding the cutoffs.

```{r}
# Plot Cook's Distance vs observation index.
PlotCookD(m2b)

# How many cases have high Cook's Distances (D > cutoff)?
summary(cooks.distance(m2b))
addmargins(table(cooks.distance(m2b) > CookDco(m2b)))

# Plot Leverage (hat) values vs. observation index.
PlotHat(m2b)

# Plot Standardized Pearson Residuals vs Leverage, w/ Cook's D contours.
plot(m2b, which = 5, id.n = 10, cex.id = .6, 
     cook.levels = round(CookDco(m2b), digits = 3))
abline(v = hatco(m2b), lty = 2, col = "blue")
text(x = hatco(m2b), y = 5.7, pos = 4, col = "blue", cex = .75,
     labels = paste("Leverage cutoff >", round(hatco(m2b), digits = 3)))

# How many cases have high leverage (hat > cutoff)?
summary(hatvalues(m2b))
addmargins(table(hatvalues(m2b) > hatco(m2b)))

# Identify cases w/ high leverage and high Cook's D. 
InfCases(m2b)
```

\FloatBarrier

## Graphs

\FloatBarrier

### Interaction Effect

We visualize below the level-transition-specific OPIc effect. The effect of OPIc
decreases as we go from the lowest to the highest level transition, as evidenced
by the decreasingly steep slope from the left-most to the right-most panel.

```{r}
visreg(m2b, xvar = "COPIC", by = "Level", layout = c(4, 1), jitter = TRUE,
       strip.names = TRUE, scale = "linear", 
       xlab = "Centered OPIC Score (OPIC - 5)", 
       ylab = "Log odds (Pass)")
visreg(m2b, xvar = "COPIC", by = "Level", layout = c(4, 1), jitter = TRUE,
       strip.names = TRUE, scale = "response", rug = 2,
       xlab = "Centered OPIC Score (OPIC - 5)", 
       ylab = "Conditional Pass Rate")
```

\FloatBarrier

### Conditional Pass Rates
Use ggplot to generate the graph for conditional pass rates derived from model
m2b.

```{r}
F3 <- ggplot(T32, aes(Level, Pass.Rate, group=OPIC, color=OPIC))+
      geom_line()+ coord_cartesian(ylim = c(0, 1))+
      xlab("SA Level Transition")+
      ylab("Conditional Pass Rate")+
      ggtitle(label="Conditional Pass Rates by OPIC Score and Level Transition")
direct.label(F3, "first.polygons")
```

\FloatBarrier

### Unconditional Pass Rates  
Use ggplot to generate the graph for unconditional pass rates derived from model
m2b.

```{r}
F4 <- ggplot(T32, aes(Level, Pass.URate, group=OPIC, color=OPIC))+
      geom_line()+ coord_cartesian(ylim = c(0, 1))+
      xlab("SA Level Transition")+
      ylab("Unconditional Pass Rate")+
      ggtitle(label="Unconditional Pass Rates by OPIC Score and SA Level Transition")
direct.label(F4, "first.polygons")
```

\FloatBarrier

## Conclusion
Model 2b is better than model 2a because the Level x COPIC interaction is
significant. It has better accuracy and sensitivity than model 1a, but we need
to run another model to decide whether adding a course effect to model 2b will
improve the model. That will be done in model 3a.

\FloatBarrier

# MODEL 3A: PARALLEL COURSE + PARALLEL OPIC EFFECTS

We now test a model with a parallel course effect and a parallel OPIc
effect. We want to see if course and OPIc are still significant preditors 
of the condtional pass rates when they are simultaneously included into one 
regression model. Model 3a tells us to what extent OPIc and Course each predicts
the conditional pass rates, after controlling for the other predictor's effect. 

```{r}
T34 <- summary(m3a)
set.caption("Model 3a Summary")
pander(T34, digits = c(5, 5, 4, 3), round = c(4, 4, 2, Inf), justify = "right",
       caption = "Model 3a Summary")
```

Post-process the raw model output to get additional information, including 
s-values, BFBs, and posterior probabilities of H1.

```{r}
format(convertp(T34$coefficients[,"Pr(>|z|)"], digits = 2), digits = 3)
```

\FloatBarrier

## Sequential Tests (Type I SS)
Each row in Table 36 tests the significance of unique additional variance
explained by the term on that line after controlling for all previously entered
terms. Significant results mean adding that term improved the model. 

```{r}
T36 <- anova(m3a, test = "Chisq")
T36 <- cbind(T36, convertp(T36[,"Pr(>Chi)"]))
pander(T36, justify = "right", split.tables = Inf, 
       digits = c(5, 5, 5, 8, 3, 3, 4, 3, 3), 
       round = c(0, 3, 0, 3, Inf, 2, 2, 2), 
       caption = "Model 3a Sequential Tests (Type I SS): Analysis of Deviance")
```

Note that Course entered the model last, and it was not significant. This 
indicated that Course did not explain a significant amount of the variance 
in the conditional pass rates after controlling for COPIC.

\FloatBarrier

## Simultaneous Tests of Main/Interaction Effects via LRT (Type III SS)
The simultaneous tests in Table 37 are the effects of the indicated terms after
controlling for all other terms in the model. They are only computed for terms
that are not part of a higher-order interaction because it makes no sense to
test for a main effect when the variable is involved in an interaction. These
should be functionally equivalent to the results you get with *anova()* when you
feed it a pair of nested models that differ only in that one model includes a
term that is absent from the other model. They are likelihood ratio tests
(LRTs).

```{r}
T37 <- drop1(m3a, test = "Chisq")   
T37 <- cbind(T37, convertp(T37[,"Pr(>Chi)"]))
pander(T37, justify = "right", split.tables = Inf, 
       digits = c(5, 8, 5, 8, 3, 4, 3, 3), 
       round = c(0, 3, 0, 3, Inf, 2, 2, 2),
       caption = "Model 3a Simultaneous Tests (Type III SS)")
```

Results of the simultaneous tests are consistent with those of the sequential 
tests in showing that Course was no longer a significant predictor of the 
conditional pass rates after controlling for COPIC.

The fact that the course effect is not significant in Table 37 tells us that we
can really revert back to model 2a because Course doesn't have an effect after
controlling for level and COPIC.

\FloatBarrier

## Conditional and Unconditional Pass Rates
We omit calculating these rates because model 3a is not better than model 2a.

\FloatBarrier

## Odds-Ratio for COPIC Effect 
We omit calculating odds-ratios because model 3a is not better than model 2a.

\FloatBarrier

## Assessing Goodness of Fit, Discrimination, and Calibration
Since we are running the continuation-ration model as a logistic regression, all
the stadard methods for assessing goodness of fit for logistic regresion models
apply.

\FloatBarrier

### Hosmer-Lemeshow goodness of fit test. 
The null hypothesis for the Hosmer-Lemeshow test is that the data fit the model,
so a significant result is actually undesirable because it says that the data
don't fit the model. This is a measure of calibration (Fenlon et al., 2018), 
which can also be examined with a calibration plot like the one shown below.

The calibration plot uses the same set of bins used for the HLT. We used a
binning method that creates bins by grouping the predicted probabilities into
bins with a minimum width of 0.10 because that way all the values grouped
together are quite similar.

```{r}
HLT.m3a <- HLfit(m3a, bin.method = "prob.bins", min.prob.interval = 0.1, 
                 main = "Calibration Plot", 
                 xlab = "Predicted Probability (minimum bin interval = 0.10)",
                 ylab = "Observed Probability")
HLT.m3a
```

```{r}
# Get the s-value, BFB, and posterior probability of H1. 
convertp(p = HLT.m3a$p.value, digits = 2)
```

\FloatBarrier

### Classification 
Here we want to examine the classification table, plus the overall accuracy,
sensitivity, and specificity estimates.

```{r}
TAData$pred.m3a <- predict(m3a, type = "response")
set.seed(3419) # For reproducibility of bootstrap estimates.
roc.m3a <- roc(pass ~ pred.m3a, data = TAData, ci = TRUE, direction = "<",
               ci.method = "bootstrap", boot.n = 10000)
round(lrcm(roc.m3a, seed = 6342), digits = 3) 
```

\FloatBarrier

### Brier Score
The unscaled Brier score is an overall measure of accuracy suitable for logistic
regression models with binary outcomes (Fenlon et al., 2018). It is the average
prediction error (Steyerberg et al., 2001; Steyerberg et al., 2010). The scaled 
Brier score adjusts for unscaled version to have a range of from 0 to 1, thereby
making it similar to an $R^{2}$ statistic. We report the scaled Brier score 
below.  

```{r}
brier(m3a, digits = 2)
```

\FloatBarrier

### Area Under the Curve (AUC)
The area under the receiver operating characteristic curve (AUC) measures the
model's ability to discriminate those who pass from those who fail to pass the
level transitions. It can range from 0.50 to 1.00; values of 0.50--0.69 are
poor, 0.70-0.79 are acceptable, .80-.89 are excellent, and $\geq$ 0.90 are
outstanding.

```{r}
print(roc.m3a)
plot.roc(roc.m3a, print.auc = TRUE, main = "ROC Curve, Model 3a", 
         print.thres = "best")
```

Next we compare the ROC curves for models m3a and m2a using a bootstrap test for
the difference between the AUC values (Robin et al., 2011). 

```{r}
set.seed(1346) # For reproducible bootstrap estimates.
roct.m2a.m3a <- roc.test(roc.m2a, roc.m3a, method = "bootstrap", boot.n = 10000)
roct.m2a.m3a
```

```{r}
# Get the s-value, BFB, and posterior probability of H1. 
convertp(p = roct.m2a.m3a$p.value, digits = 2)
```

\FloatBarrier

### $R^{2}$ Measures
Pseudo-$R^{2}$ ($R^{2}_p$) is the squared Pearson correlation between observed
and predicted values, as suggested in Hosmer, Lemeshow, & Sturdivant (2013, p.
182). Meanwhile, $R^{2}_{Dev}$ is a measure based on deviance residuals (Fox,
1997, p. 451; Cameron & Windmeijer, 1997).

```{r}
c(PsedoR2 = pseudoR2(m3a, digits = 2), R2Dev = R2Dev(m3a, digits = 2))
```

\FloatBarrier

## Diagnostics
We omit diagnostics because model 3a is not better than model 2a.

\FloatBarrier

## Graphs
We omit graphs because model 3a is not better than model 2a. 

\FloatBarrier

## Conclusion
What distinguished this model from model 2a was inclusion of the parallel course
effect, but the Type III LRT for that effect is not-significant. We will report
model 2a over this model on the basis of parsimony.


# MODEL 3B: PARALLEL COURSE + NON-PARALLEL OPIC EFFECTS

We finally test a model with a parallel course effect and a non-parallel OPIc
effect. We add the course effect into the model after the interaction so that a
sequential test will examine whether it adds any value beyond the non-parallel
OPIC effect.

```{r}
T38 <- summary(m3b)
set.caption("Model 3a Summary")
pander(T38, digits = c(5, 5, 4, 3), round = c(4, 4, 2, Inf), justify = "right",
       caption = "Model 3a Summary")
```

Post-process the raw model output to get additional information, including 
s-values, BFBs, and posterior probabilities of H1.

```{r}
format(convertp(T38$coefficients[,"Pr(>|z|)"], digits = 2), digits = 3)
```

\FloatBarrier

## Sequential Tests (Type I SS)
Each row in Table 40 tests the significance of unique additional variance
explained by the term on that line after controlling for all previously entered
terms. Significant results mean adding that term improved the model. Note that 
course actually entered the model last. 

```{r}
T40 <- anova(m3b, test = "Chisq")
T40 <- cbind(T40, convertp(T40[,"Pr(>Chi)"]))
pander(T40, justify = "right", split.tables = Inf, 
       digits = c(5, 5, 5, 8, 3, 3, 4, 3, 3), 
       round = c(0, 3, 0, 3, Inf, 2, 2, 2), 
       caption = "Model 3a Sequential Tests (Type I SS): Analysis of Deviance")
```

\FloatBarrier

## Simultaneous Tests of Main/Interaction Effects via LRT (Type III SS)
The simultaneous tests in Table 41 are the effects of the indicated terms after
controlling for all other terms in the model. They are only computed for terms
that are not part of a higher-order interaction because it makes no sense to
test for a main effect when the variable is involved in an interaction. These
should be functionally equivalent to the results you get with *anova()* when you
feed it a pair of nested models that differ only in that one model includes a
term that is absent from the other model. They are likelihood ratio tests
(LRTs).

```{r}
T41 <- drop1(m3b, test = "Chisq")   
T41 <- cbind(T41, convertp(T41[,"Pr(>Chi)"]))
pander(T41, justify = "right", split.tables = Inf, 
       digits = c(5, 8, 5, 8, 3, 4, 3, 3), 
       round = c(0, 3, 0, 3, Inf, 2, 2, 2),
       caption = "Model 3a Simultaneous Tests (Type III SS)")
```

The fact that the course effect is not significant in Table 41 tells us that we
can really revert back to model 2b because course doesn't have an effect after
controlling for level, COPIC, and their interaction.

\FloatBarrier

## Conditional and Unconditional Pass Rates
We omit calculating these rates because model 3b is not better than model 2b.

\FloatBarrier

## Odds-Ratio for COPIC Effect 
We omit calculating odds-ratios because model 3b is not better than model 2b.

\FloatBarrier

## Assessing Goodness of Fit, Discrimination, and Calibration
Since we are running the continuation-ration model as a logistic regression, all
the stadard methods for assessing goodness of fit for logistic regresion models
apply.

\FloatBarrier

### Hosmer-Lemeshow goodness of fit test. 
The null hypothesis for the Hosmer-Lemeshow test is that the data fit the model,
so a significant result is actually undesirable because it says that the data
don't fit the model. This is a measure of calibration (Fenlon et al., 2018), 
which can also be examined with a calibration plot like the one shown below.

The calibration plot uses the same set of bins used for the HLT. We used a
binning method that creates bins by grouping the predicted probabilities into
bins with a minimum width of 0.10 because that way all the values grouped
together are quite similar.

```{r}
HLT.m3b <- HLfit(m3b, bin.method = "prob.bins", min.prob.interval = 0.1, 
                 main = "Calibration Plot", 
                 xlab = "Predicted Probability (minimum bin interval = 0.10)",
                 ylab = "Observed Probability")
HLT.m3b
```

```{r}
# Get the s-value, BFB, and posterior probability of H1. 
convertp(p = HLT.m3b$p.value, digits = 2)
```

\FloatBarrier

### Classification 
Here we want to examine the classification table, plus the overall accuracy,
sensitivity, and specificity estimates.

```{r}
TAData$pred.m3b <- predict(m3b, type = "response")
set.seed(9173) # For reproducibility of bootstrap estimates.
roc.m3b <- roc(pass ~ pred.m3b, data = TAData, ci = TRUE, direction = "<",
               ci.method = "bootstrap", boot.n = 10000)
round(lrcm(roc.m3b, seed = 7146), digits = 3) 
```

\FloatBarrier

### Brier Score
The unscaled Brier score is an overall measure of accuracy suitable for logistic
regression models with binary outcomes (Fenlon et al., 2018). It is the average
prediction error (Steyerberg et al., 2001; Steyerberg et al., 2010). The scaled 
Brier score adjusts for unscaled version to have a range of from 0 to 1, thereby
making it similar to an $R^{2}$ statistic. We report the scaled Brier score 
below.  

```{r}
brier(m3b, digits = 2)
```

\FloatBarrier

### Area Under the Curve (AUC)
The area under the receiver operating characteristic curve (AUC) measures the
model's ability to discriminate those who pass from those who fail to pass the
level transitions. It can range from 0.50 to 1.00; values of 0.50--0.69 are
poor, 0.70-0.79 are acceptable, .80-.89 are excellent, and $\geq$ 0.90 are
outstanding.

```{r}
print(roc.m3b)
plot.roc(roc.m3b, print.auc = TRUE, main = "ROC Curve, Model 3a", 
         print.thres = "best")
```

Next we compare the ROC curves for models m2a and m3b using a bootstrap test for
the difference between the AUC values (Robin et al., 2011). 

```{r}
set.seed(1128) # For reproducible bootstrap estimates.
roct.m2a.m3b <- roc.test(roc.m2a, roc.m3b, method = "bootstrap", boot.n = 10000)
roct.m2a.m3b
```

```{r}
# Get the s-value, BFB, and posterior probability of H1. 
convertp(p = roct.m2a.m3b$p.value, digits = 2)
```

\FloatBarrier

### $R^{2}$ Measures
Pseudo-$R^{2}$ ($R^{2}_p$) is the squared Pearson correlation between observed
and predicted values, as suggested in Hosmer, Lemeshow, & Sturdivant (2013, p.
182). Meanwhile, $R^{2}_{Dev}$ is a measure based on deviance residuals (Fox,
1997, p. 451; Cameron & Windmeijer, 1997).

```{r}
c(PsedoR2 = pseudoR2(m3b, digits = 2), R2Dev = R2Dev(m3b, digits = 2))
```

\FloatBarrier

## Diagnostics
We omit diagnostics because model 3b is not better than model 2b.

\FloatBarrier

## Graphs
We omit graphs because model 3b is not better than model 2b. 

\FloatBarrier

## Conclusion
What distinguished this model from model 2b was inclusion of the parallel course
effect, but the Type III LRT for that effect is not-significant. We will report
model 2b over this model on the basis of parsimony.

\FloatBarrier

# References
Benjamin, D. J., & Berger, J. O. (2019). Three recommendations for improving 
the use of p-values. *The American Statistician, 73*(Supplement 1), 186-191. 
https://doi.org/10.1080/00031305.2018.1543135

Cameron, A. C., & Windmeijer, F. A. G. (1997). An R-squared measure of goodness
of fit for some common nonlinear regression models. *Journal of Econometrics,
77*(2), 329-342. https://doi.org/10.1016/S0304-4076(96)01818-0

Colquhoun, D. (2019). The false positive risk: A proposal concerning what to 
do about p-values. *The American Statistician, 73*(Supplement 1), 192-201. 
https://doi-org/10.1080/00031305.2018.1529622

Fenlon, C., OGrady, L., Doherty, M. L., & Dunnion, J. (2018). A discussion of 
calibration techniques for evaluating binary and categorical predictive models.
*Preventive Veterinary Medicine, 149*, 107-114. 
https://doi.org/10.1016/j.prevetmed.2017.11.018

Fox, J. (1997). *Applied regression analysis, linear models, and related
methods*. Thousand Oaks, CA: Sage Publications.

Greenland, S. (2019). Valid p-values behave exactly as they should: Some 
misleading criticisms of p-values and their resolution with s-values. 
*The American Statistician, 73*(Supplement 1), 106-114. 
https://doi.org/10.1080/00031305.2018.1529625

Hosmer, D. W., Lemeshow, S., & Sturdivant, R. X. (2013). *Applied logistic
regression* (3rd ed.). Hoboken, NJ: John Wiley & Sons, Inc.

Perkins, N. J., & Schisterman, E. F. (2006). The inconsistency of optimal
cutpoints obtained using two criteria based on the receiver operating
characteristic curve. *American Journal of Epidemiology, 163*(7), 670-675.
https://doi-org/10.1093/aje/kwj063

Raykov, T., & Marcoulides, G. A. (2011). *Introduction to psychometric theory*.
New York, NY: Routledge.

Robin, X., Turck, N., Hainard, A., Tiberti, N., Lisacek, F., Sanchez, J.-C., & 
Mller, M. (2011). pROC: an open-source package for R and S+ to analyze and 
compare ROC curves. *BMC Bioinformatics, 12*, 77. 
https://doi.org/10.1186/1471-2105-12-77

Steyerberg, E. W., Harrell Jr., F. E., Borsboom, G. J. J. M., Eijkemans, M. J. 
C., Vergouwe, Y., & Habbema, J. D. F. (2001). Internal validation of predictive 
models: Efficiency of some procedures for logistic regression analysis. Journal 
of Clinical Epidemiology, 54(8), 774-781. 
https://doi.org/10.1016/S0895-4356(01)00341-9

Steyerberg, E. W., Vickers, A. J., Cook, N. R., Gerds, T., Gonen, M., 
Obuchowski, N. A., . . . Kattan, M. W. (2010). Assessing the performance of 
prediction models : A framework for traditional and novel measures. 
Epidemiology, 21(1), 128-138. http://doi.org/10.1097/EDE.0b013e3181c30fb2

Wasserstein, R. L., Schirm, A. L., & Lazar, N. A. (2019). Moving to a world 
beyond "p < .05". *The American Statistician, 73*(Supplement 1), 1-19. 
https://doi.org/10.1080/00031305.2019.1583913

Westfall, P. H. (1997). Multiple testing of general contrasts using logical 
constraints and correlations. *Journal of the American Statistical Association, 
92*(437), 299-306. https://doi.org/10.1080/01621459.1997.10473627

Winke, P., Zhang, X., & Pierce, S. J. (2022). 
A closer look at a marginalized test method: Self-assessment as a measure of 
speaking proficiency [Manuscript accepted for publication]. 
*Studies in Second Language Acquisition*.

Winke, P., Pierce, S. J., & Zhang, X. (2018, October). Self-assessment works! 
Continuation-ratio models for testing course and OPIc score effects on oral 
proficiency self-assessments. Paper presented at the East Coast Organization 
of Language Testers 2018 conference, hosted by the Educational Testing Service,
Princeton, NJ. https://sites.google.com/site/ecoltaelrc/home

Youden, W. J. (1950). Index for rating diagnostic tests. *Cancer, 3*(1), 32-35.
https://doi.org/10.1002/1097-0142(1950)3%3A1%3C32%3A%3AAID-CNCR2820030106%3E3.0.CO%3B2-3

\FloatBarrier

# Software Information
We use R Markdown to enhance reproducibility. Knitting the source R Markdown 
script generates the PDF file containing explanatory text, R code, and R output.

- We used [RStudio](www.rstudio.org) `r rstudioapi::versionInfo()$version` to 
  work with R and R markdown files. 
- Our software chain looks like this:
  **Rmd file > RStudio > R > rmarkdown > knitr > md file > pandoc > tex file > TinyTeX > PDF file**.
- Source script: *`r knitr:::current_input()`*
- Output file: *`r sub(".Rmd", ".pdf", knitr:::current_input(dir = FALSE))`*
- We recommend using [TinyTeX](https://yihui.org/tinytex/) to compile LaTeX 
  files into PDF files. However, it should be viable to use 
  [MiKTeX](https://miktex.org) instead. 
- We used [pandoc](https://pandoc.org) `r pandoc_version()` for this 
  document. 

This document was generated using the following computational environment and 
dependencies: 

``` {r show-citations}
# Check and report whether we used TinyTex or other LaTeX software. 
which_latex()

# Get R and R package version numbers in use.
devtools::session_info()
```

The current Git commit details and status are:

```{r git-details, echo = TRUE, eval=TRUE}
git_report()
```
