---
title: \sffamily{\bfseries{\LARGE Self-Assessment Works! Import Data}}
geometry: "left=1.0in,right=1.0in,top=.75in,bottom=.75in"
author: "Steven J. Pierce & Xiaowan Zhang"
output:
 pdf_document:
   latex_engine: xelatex
   number_sections: true
   toc: yes
   toc_depth: 3
urlcolor: blue
header-includes:
- \usepackage{fancyhdr}
- \usepackage[yyyymmdd,hhmmss]{datetime}
- \usepackage{lastpage}
- \usepackage{amsmath,amsthm}
- \usepackage{unicode-math}
- \defaultfontfeatures{Ligatures=TeX}
- \usepackage[font={small}, margin=1cm, skip=2pt]{caption}
- \usepackage{url}
- \usepackage{floatrow} 
- \floatplacement{figure}{!ht}
- \floatplacement{table}{!ht}
- \usepackage{placeins}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{dcolumn}
- \usepackage{titling}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \pretitle{\begin{center}\LARGE\bfseries}
- \posttitle{\end{center}}
- \usepackage{lscape}
- \pagestyle{fancy}
- \lhead{Self-Assessment Works! (Spanish) Import Data}
- \rhead{\today\ \currenttime}
- \cfoot{ }
- \lfoot{\texttt{\footnotesize \detokenize{`r sub(".Rmd", ".pdf", knitr:::current_input(dir = FALSE))`}}} 
- \fancyfoot[R]{\thepage\ of \pageref*{LastPage}}
- \renewcommand{\headrulewidth}{0.4pt}
- \renewcommand{\footrulewidth}{0.4pt}
- \fancypagestyle{plain}{\pagestyle{fancy}}
---

\FloatBarrier

# Purpose
This file facilitates reproducing the results reported in our manuscript (Winke,
Zhang, & Pierce, 2021), which was based on a presentation (Winke, Pierce, &
Zhang 2018). It imports data on Spanish language students from SPSS data files 
into R, runs some initial data analyses to explore key variables, and saves out 
an `.RData` file for use by subsequent scripts. 

# Setup
Set global R chunk options (local chunk options will over-ride global options). 
The method for creating a size option that controls font size in code chunks and
their text output is based on an answer to a question posted on   
[stackoverflow.com](https://stackoverflow.com/questions/25646333/code-chunk-font-size-in-rmarkdown-with-knitr-and-latex/46526740#46526740). 

``` {r global-options, cfsize = "footnotesize"}
# Create a custom chunk hook/option for controlling font size in chunk & output.
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$cfsize != "normalsize", paste0("\n \\", options$cfsize,"\n\n", 
                                              x, "\n\n \\normalsize"), x)
  })

# Global chunk options (over-ridden by local chunk options)
knitr::opts_chunk$set(include  = TRUE, echo = TRUE, error = TRUE, 
                      message = TRUE, warning = TRUE, fig.pos = "!ht", 
                      cfsize = "footnotesize")

# Declare location of this script relative to the project root directory.
here::i_am(path = "inst/SAW_Paper_Import_Explore_Data.Rmd")
```

Some of the functions we use below come from a personal R package called
[*piercer*](https://github.com/sjpierce/piercer) that is available online from 
[GitHub](https://github.com). Please read and follow the installation 
instructions for *piercer* at https://github.com/sjpierce/piercer
before knitting this script, otherwise the next chunk will not load that 
package and various parts of this script will not work.  

Load contributed R packages that we need to get additional functions. 

``` {r load-packages}
library(here)              # for here()
library(devtools)          # for session_info()
# Set package options. 
# options(knitr.kable.NA = '0.00')
library(rmarkdown)         # for render(), pandoc_version(). 
library(knitr)             # for kable()
library(tidyr)             # for unite()
library(dplyr)             # for filter(), select(), etc.
library(sjlabelled)       # For set_label(), get_label()
library(haven)            # for read_spss()
# Set package options. 
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)        # for kable_styling(), add_header_above(), 
                           # column_spec(), collapse_rows(), and landscape()
library(piercer)           # for git_report(), which_latex()
library(SAWpaper)          # for package version number via session_info()
```

\FloatBarrier

# Import Data
This section of the script imports the required data from SPSS data files into
R data frames and saves out an .RData file. 

Please see the `README.Rmd` or `README.md` files in this *SAWpaper* package for
details on obtaining the SPSS data files referenced by the code below. The code
chunks show the relative file paths for where you need to put the files in order
for the script to work as intended. 

\FloatBarrier

## Student-Level Data (for Correlation Analyses)
The file `Students_Data.sav` has the cleaned self-assessment data on a set of
Spanish students, with one row per student. We prepared an SPSS file to simplify
archiving and attaching meta-data such as variable and value labels, level of
measurement, and so on. The chunk below reads that file into a data frame called
`SData`, then converts a couple variables to proper factors.

```{r import-student-data}
SData <- read_sav(here::here("inst/extdata/Students_Data.sav"), user_na = TRUE) %>% 
  mutate(Gender = as_factor(Gender),
         Course = as_factor(Course),
         LevelF = as_factor(Level, levels = "values"))
```

Next, we create a smaller data frame called `VSData` containing only students 
who had valid OPIC scores (i.e., that were not above or below range or 
unratable). The latter is the set we used for most analyses.  

```{r subset-student-data}
VSData <- SData %>% filter(OPICA >= 1)
```

The `SData` data frame contains `r nrow(SData)` rows and `r ncol(SData)` 
variables, while the `VSData` data frame contains `r nrow(VSData)` rows and 
`r ncol(VSData)` variables. 

\FloatBarrier

## Testlet-Level Data (for Continuation-Rato Modeling)
The file `Testlet_Attempts_Data.sav` contains data about the same set of
students (those identified in `Students_Data.sav`) that has been expanded to one
row per student per self-assessment testlet attempted. This data structure is
conceptually similar to a person-period file for longitudinal analysis. The
chunk below reads that file into a data frame called `TAData`, then converts a
couple variables to proper factors. 

```{r import-crm-testlet-data}
TAData <- read_sav(here::here("inst/extdata/Testlet_Attempts_Data.sav"), 
                   user_na = TRUE) %>% 
  mutate(Gender = as_factor(Gender),
         Course = as_factor(Course),
         Testlet = as_factor(Testlet, levels = "values"),
         MaxLevel = as_factor(MaxLevel, levels = "values"))
```

Next, we create a smaller data frame called `VTAData` containing only data students 
who had valid OPIC scores (i.e., that were not above or below range or 
unratable). The latter is the set we used for most analyses. We need the 
`VTAData` data frame specifically to run continuation-ratio models. Most other 
analyses are performed based on the `VSData` data frame.

```{r subset-crm-testlet-data}
VTAData <- TAData %>% filter(OPICA >= 1)
```

The `TAData` data frame contains `r nrow(TAData)` rows and `r ncol(TAData)` 
variables, while the `VTAData` data frame contains `r nrow(VTAData)` rows and 
`r ncol(VTAData)` variables. 

\FloatBarrier

### Contrast Coding 
Tables \ref{tab:Testlet-contrasts} and \ref{tab:Course-contrasts} below show the 
default contrast coding associated with key factor variables in `TAData` so 
readers know how to interpret model coefficients later. The default for factors 
should be dummy coding (treatment contrasts) with the first level set as the
reference group. 

```{r Testlet-contrasts}
# Table caption and footnote text. 
TCap <- "Default Contrast Dummy Coding for Testlet"
FN   <- "The reference value is the one always coded as zero."

# Generate a table showing the default contrast coding. 
contrasts(TAData$Testlet) %>% 
  kable(format = "latex", booktabs = TRUE, row.names = TRUE, caption = TCap) %>% 
  kable_styling(latex_options = c("repeat_header")) %>% 
  add_header_above(c("Testlet", "Contrast Coding" = 3)) %>% 
  column_spec(1:3, width = ".5cm") %>% 
  footnote(general = FN, general_title = "Note: ", footnote_as_chunk = TRUE,
           threeparttable = TRUE)
```

```{r Course-contrasts}
# Table caption and footnote text. 
TCap <- "Default Contrast Dummy Coding for Course"
FN   <- "The reference value is the one always coded as zero."

# Generate a table showing the default contrast coding. 
contrasts(TAData$Course) %>% 
  kable(format = "latex", booktabs = TRUE, row.names = TRUE, caption = TCap) %>% 
  kable_styling(latex_options = c("repeat_header")) %>% 
  add_header_above(c("Course", "Contrast Coding" = 3)) %>% 
  column_spec(1:3, width = ".5cm") %>% 
  footnote(general = FN, general_title = "Note: ", footnote_as_chunk = TRUE,
           threeparttable = TRUE)
```

\FloatBarrier

### OPIC Scores
We treat OPIc speaking test proficiency scores as a continuous covariate that
ranges from Novice-low to Superior. Table \ref{tab:OPIC-scores} shows the coding
for the `OPICA`, `OPICV`, and `COPIC` variables in `SData`. The `OPICA` variable 
contains all possible OPIc ratings, including those for individuals who were 
identified as above or below range or unratable (AR, BR, or UR) in numeric form 
and has value labels attached that show the meaning of scores. Meanwhile, 
`OPICV` is a transformation of `OPICA` that replaced the AR, BR, and UR values
with missing values. To make model coefficients more interpretable, we computed
centered OPIc scores from `OPICV` and stored them in a variable named `COPIC`. A 
score of zero in `COPIC` represents the Intermediate-mid level on the OPIc 
scale.

```{r OPIC-scores}
Total.N <- nrow(SData)
Valid.N <- nrow(VSData)

# Table caption and footnote text. 
TCap <- paste0("OPIc Score Variables: Coding and Frequency Distribution (",
              "Total N = ", Total.N, ", Valid N = ", Valid.N, ")")
FN   <- paste("COPIC is a centered form of OPICV (COPIC = OPICV - 5) used to",
              "make model intercept terms more interpretable.",
              "OPICA, numeric ACTFL Oral Proficiency Interview – computer score",
              "OPICV, valid numeric ACTFL Oral Proficiency Interview – computer score.")

data.frame(OPIC_Label = get_labels(SData$OPICA), 
           OPICA = c(as.numeric(sort(unique(SData$OPICA))), 10), 
           OPICV = c(NA, NA, NA, as.numeric(sort(unique(SData$OPICV))), 10), 
           COPIC = c(NA, NA, NA, as.numeric(sort(unique(SData$COPIC))), 5),
           Frequency = as.numeric(c(xtabs(~OPICA, addNA = TRUE, data = SData), 
                                    0))) %>% 
  mutate(Percent = 100*Frequency/Total.N,
         Valid = if_else(OPICA >= 1, true = 100*Frequency/Valid.N, false = 0)) %>% 
  kable(format = "latex", booktabs = TRUE, digits = 2, 
        col.names = c("Score Label", "OPICA", "OPICV", "COPIC", "Frequency",
                      "Percent", "Valid Percent"), 
        caption = TCap) %>% 
  kable_styling(latex_options = c("repeat_header")) %>% 
  column_spec(1, width = "4cm") %>% 
  column_spec(2:4, width = "1.5cm") %>% 
  footnote(general = FN, general_title = "Note: ", footnote_as_chunk = TRUE,
           threeparttable = TRUE)
```

\FloatBarrier

# Explore Data 
This section of the script performs some initial data analyses that explore 
variable distributions and provide descriptive summaries we may use in the
manuscript. 

\FloatBarrier

## Student-Level Data (All Students)
We start by examining the student-level data in `SData`. Table 
\ref{tab:Table-LevelF} below summarizes the univariate distribution for 
maximum self-assessment level achieved by the students. 

```{r Table-LevelF}
# Table caption and footnote text. 
TCap <- "Frequency Table for Maximum Self-Assesment Level Achieved (All Students)"
FN   <- paste("Remaining N is the number of students who reached a level above",
              "the one associated with the current row in the table.")

# Frequency table for maximum level achieved (All Students). 
TLevelF <- SData %>% 
  group_by(LevelF) %>% 
  summarise(N = n(),
            Pct = 100*N/Total.N) %>% 
  ungroup() %>% 
  mutate(CumN = cumsum(N),
         CumPct = cumsum(Pct),
         RemainN = sum(N) - CumN,
         RemainPct = 100 - CumPct)

# Display formatted table.
kable(TLevelF, format = "latex", booktabs = TRUE, digits = 0, 
      col.names = c("Level", "N", "%", "Cumulative N", "Cumulative %", 
                    "Remaining N", "Remaining %"),
      caption = TCap) %>% 
  footnote(general = FN, general_title = "Note: ", footnote_as_chunk = TRUE,
           threeparttable = TRUE)
```

We now examine the relationships among Level, Course, and OPIC by creating 
crosstabulations. Table \ref{tab:Table-LevelF-Course} shows the two-way 
contingency table for Level and Course. 

```{r Table-LevelF-Course}
# Table caption and footnote text. 
TCap <- paste("Crosstab Maximum Self-Assessment Level by Course, With Marginal", 
              "Sums (All Students)")
FN   <- paste("Course, type of Spanish course the student was enrolled in;",
              "Level, maximum self-assessment level achieved by the student.")

# Examine the relationship between SA level and course level. 
addmargins(xtabs(~LevelF + Course, data = SData)) %>% 
  kable(format = "latex", booktabs = TRUE, caption = TCap) %>%
  kable_styling(latex_options = c("repeat_header")) %>% 
  add_header_above(c("Level", "Course" = 4, " ")) %>% 
  row_spec(row = 6, italic = TRUE) %>%
  column_spec(column = 6, italic = TRUE) %>% 
  footnote(general = FN, general_title = "Note: ", footnote_as_chunk = TRUE,
           threeparttable = TRUE)
```

Table \ref{tab:Table-LevelF-OPICA-Course} summarizes the cross-tabulations 
between OPIc scores and two other variables: maximum self-assessment level
achieved and course. 

```{r Table-LevelF-OPICA-Course}
# Table caption and footnote text. 
TCap <- paste("Crosstabs of Maximum Self-Assessment Level and Course by OPIc", 
              "Score, With Marginal  (All Students)")
FN   <- paste("Course, type of Spanish course the student was enrolled in;",
              "Level, maximum self-assessment level achieved by the student;",
              "OPIc, ACTFL Oral Proficiency Interview – computer.",
              "The invalid OPIc scores are: -2 = above range (AR), -1 = below",
              "range (BR), and 0 = unratable (UR).")

# Combine two crosstab outputs into a single table: 
# Examine the relationship between OPIc scores and SA level.
# Examine the relationship between OPIc scores and course level.
rbind(addmargins(xtabs(~LevelF + OPICA, addNA = TRUE, data = SData)),
      addmargins(xtabs(~Course + OPICA, addNA = TRUE, data = SData))) %>% 
  kable(format = "latex", booktabs = TRUE,  
      caption = TCap) %>% 
  kable_styling(latex_options = c("repeat_header")) %>% 
  add_header_above(header = c(" ", "Invalid" = 3, "Valid" = 9, " ")) %>% 
  add_header_above(header = c(" ", "OPIc Score" = 12, " ")) %>% 
  group_rows(group_label = "Level", bold = FALSE, start_row = 1, 
             end_row = 6) %>% 
  group_rows(group_label = "Course", bold = FALSE, start_row = 7, 
             end_row = 11) %>% 
  row_spec(row = c(6, 11), italic = TRUE) %>% 
  column_spec(column = 14, italic = TRUE) %>% 
  footnote(general = FN, general_title = "Note: ", footnote_as_chunk = TRUE,
           threeparttable = TRUE)
```

\FloatBarrier

## Student-Level Data (Students with Valid OPIc Scores)
We start by examining the student-level data in `VSData` because those are the 
data we used in most of the analyses reported in our published manuscript. Table 
\ref{tab:Table-LevelFV} below summarizes the univariate distribution for maximum 
self-assessment level achieved by the students who had valid OPIC scores. 

```{r Table-LevelFV}
# Table caption and footnote text. 
TCap <- paste("Frequency Table for Maximum Self-Assesment Level Achieved",
              "(Students with Valid OPIc Scores)")
FN   <- paste("Only students with valid OPIc scores (OPICA > 0) were included.",
              "Remaining N is the number of students who reached a level above",
              "the one associated with the current row in the table.")

# Frequency table for maximum level achieved (Students with Valid OPIC). 
TLevelFV <- VSData %>% 
  group_by(LevelF) %>% 
  summarise(N = n(),
            Pct = 100*N/Valid.N) %>% 
  ungroup() %>% 
  mutate(CumN = cumsum(N),
         CumPct = cumsum(Pct),
         RemainN = sum(N) - CumN,
         RemainPct = 100 - CumPct)

# Extract N and percent of students with self-assessed level > 1 (i.e., >= 2). 
N_SA_GE_2   <- TLevelFV$RemainN[1]
Pct_SA_GE_2 <- TLevelFV$RemainPct[1]

# Extract N and percent of students with self-assessed level > 2 (i.e., >= 3). 
N_SA_GE_3   <- TLevelFV$RemainN[2]
Pct_SA_GE_3 <- TLevelFV$RemainPct[2]


# Display formatted table.
kable(TLevelFV, format = "latex", booktabs = TRUE, digits = 0, 
      col.names = c("Level", "N", "%", "Cumulative N", "Cumulative %", 
                    "Remaining N", "Remaining %"),
      caption = TCap) %>% 
  footnote(general = FN, general_title = "Note: ", footnote_as_chunk = TRUE,
           threeparttable = TRUE)
```

Only $N = `r N_SA_GE_2`$ (`r round(Pct_SA_GE_2, digits = 0)`%) of the students 
reached level 2 or higher on the self-assessment. Only $N = `r N_SA_GE_3`$ 
(`r round(Pct_SA_GE_3, digits = 0)`%) students reached level 3 or higher on the 
self-assessment. 

We now examine the relationships among Level, Course, and OPIC by creating 
crosstabulations. Table \ref{tab:Table-LevelF-CourseV} shows the two-way 
contingency table for Level and Course. 

```{r Table-LevelF-CourseV}
# Table caption and footnote text. 
TCap <- paste("Crosstab Maximum Self-Assessment Level by Course, With Marginal",
              "Sums (Students with Valid OPIc Scores)")
FN   <- paste("Only students with valid OPIc scores (OPICA > 0) were included.",
              "Course, type of Spanish course the student was enrolled in;",
              "Level, maximum self-assessment level achieved by the student.")

# Examine the relationship between SA level and course level. 
addmargins(xtabs(~LevelF + Course, data = VSData)) %>% 
  kable(format = "latex", booktabs = TRUE, caption = TCap) %>%
  kable_styling(latex_options = c("repeat_header")) %>% 
  add_header_above(c("Level", "Course" = 4, " ")) %>% 
  row_spec(row = 6, italic = TRUE) %>%
  column_spec(column = 6, italic = TRUE) %>% 
  footnote(general = FN, general_title = "Note: ", footnote_as_chunk = TRUE,
           threeparttable = TRUE)
```

Table \ref{tab:Table-LevelF-OPICV-CourseV} summarizes the cross-tabulations 
between OPIc scores and two other variables: maximum self-assessment level
achieved and course. 

```{r Table-LevelF-OPICV-CourseV}
# Table caption and footnote text. 
TCap <- paste("Crosstabs of Maximum Self-Assessment Level and Course by OPIc", 
              "Score, With Marginal Sums (Students with Valid OPIc Scores)")
FN   <- paste("Only students with valid OPIc scores (OPICA > 0) were included.",
              "Course, type of Spanish course the student was enrolled in;",
              "Level, maximum self-assessment level achieved by the student;",
              "OPIc, ACTFL Oral Proficiency Interview – computer.")

# Combine two crosstab outputs into a single table: 
# Examine the relationship between OPIc scores and SA level.
# Examine the relationship between OPIc scores and course level.
rbind(addmargins(xtabs(~LevelF + OPICV, data = VSData)),
      addmargins(xtabs(~Course + OPICV, data = VSData))) %>% 
  kable(format = "latex", booktabs = TRUE,  
      caption = TCap) %>% 
  kable_styling(latex_options = c("repeat_header")) %>% 
  add_header_above(header = c(" ", "OPIc Score" = 9, " ")) %>% 
  group_rows(group_label = "Level", bold = FALSE, start_row = 1, 
             end_row = 6) %>% 
  group_rows(group_label = "Course", bold = FALSE, start_row = 7, 
             end_row = 11) %>% 
  row_spec(row = c(6, 11), italic = TRUE) %>% 
  column_spec(column = 11, italic = TRUE) %>% 
  footnote(general = FN, general_title = "Note: ", footnote_as_chunk = TRUE,
           threeparttable = TRUE)
```

\FloatBarrier

## Testlet-Level Data
Now we examine some additional crosstabulations based on the testlet-level data
frame `VTAData`. Note that the Testlet variable in `VTAData` records the level
transition testlet used to determine whether a student could move on from one 
self-assessment level to the next. Thus, Testlet = 1 pertains to the attempt to 
transition from level 1 to level 2, Testlet = 2 pertains to the transition from 
level 2 to level 3, and so on.  

Table \ref{tab:Table-Testlet-Pass} shows the contingency table between Testlet 
and Pass, which we can use to get crude estimates of conditional and 
unconditional pass rates (albeit without confidence intervals). The conditional
pass rate uses a denominator comprised of only students who took the given
testlet, whereas the the unconditional pass rate uses the total number of 
students in the sample as the denominator.

```{r Table-Testlet-Pass}
# Table caption and footnote text. 
TCap <- paste("Crosstab Self-Assessment Level Transition Testlet by Pass",
              "(Students with Valid OPIc Scores)")
FN   <- paste("Only students with valid OPIc scores (OPICA > 0) were included.",
              "CPR, conditional pass rate;",
              "Pass, whether student passed the testlet (0 = No, 1 = Yes);",
              "Testlet, level transition testlet attempted by student;",
              "UPR, unconditional pass rate.")

# Crosstab to get a look at how often people pass each level transition.  
addmargins(xtabs(~Testlet + Pass, data = VTAData)) %>% 
  as_tibble(.) %>% 
  pivot_wider(names_from = Pass, values_from = n) %>% 
  mutate(CPR = case_when(Testlet != "Sum" ~ `1`/Sum),
         UPR = case_when(Testlet != "Sum" ~ `1`/n_distinct(VTAData$StudentID))) %>% 
  kable(format = "latex", booktabs = TRUE, digits = 3, caption = TCap) %>%
  kable_styling(latex_options = c("repeat_header")) %>% 
  add_header_above(header = c(" ", "Pass" = 2, " ", " ", " ")) %>% 
  add_header_above(header = c(" ", "Counts" = 3, "Pass Rates" = 2)) %>% 
  row_spec(row = 5, italic = TRUE) %>%
  column_spec(column = 4, italic = TRUE) %>% 
  footnote(general = FN, general_title = "Note: ", footnote_as_chunk = TRUE,
           threeparttable = TRUE)
```

Table \ref{tab:Table-Testlet-Course} shows the contingency table between Testlet
and Course. 

```{r Table-Testlet-Course}
# Table caption and footnote text. 
TCap <- paste("Crosstab Self-Assessment Level Transition Testlet by Course", 
              "(Students with Valid OPIc Scores)")
FN   <- paste("Only students with valid OPIc scores (OPICA > 0) were included.",
              "Course, type of Spanish course the student was enrolled in;",
              "Testlet, level transition testlet attempted by student.")

# Examine the relationship between SA level transition and course level.
addmargins(xtabs(~Testlet + Course, data = VTAData)) %>% 
  kable(format = "latex", booktabs = TRUE, caption = TCap) %>%
  kable_styling(latex_options = c("repeat_header")) %>% 
  add_header_above(header = c("Testlet ", "Course" = 4, " ")) %>% 
  row_spec(row = 5, italic = TRUE) %>%
  column_spec(column = 6, italic = TRUE) %>% 
  footnote(general = FN, general_title = "Note: ", footnote_as_chunk = TRUE,
           threeparttable = TRUE)
```

Table \ref{tab:Table-Testlet-COPIC} summarizes the cross-tabulations 
between Testlet and centered OPIc scores. 

```{r Table-Testlet-COPIC}
# Table caption and footnote text. 
TCap <- paste("Crosstab Self-Assessment Level Transition Testlet by Centered", 
              "OPIc Score (Students with Valid OPIc Scores)")
FN   <- paste("Only students with valid OPIc scores (OPICA > 0) were included.",
              "COPIC is a centered form of OPICN (COPIC = OPICN - 5) used to",
              "make model intercept terms more interpretable.",
              "OPICN, numeric ACTFL Oral Proficiency Interview – computer score;",
              "Testlet, level transition testlet attempted by student.")

# Examine the relationship between centered OPIc scores and SA level transition.
addmargins(xtabs(~Testlet + COPIC, data = VTAData)) %>% 
  kable(format = "latex", booktabs = TRUE, caption = TCap) %>%
  kable_styling(latex_options = c("repeat_header")) %>% 
  add_header_above(header = c("Testlet", "COPIC" = 9, 
                              " ")) %>% 
  row_spec(row = 5, italic = TRUE) %>%
  column_spec(column = 11, italic = TRUE) %>% 
  footnote(general = FN, general_title = "Note: ", footnote_as_chunk = TRUE,
           threeparttable = TRUE)
```

\FloatBarrier

# Save Data 
The chunks below save the `SData` and `TAData` data frames to an `.RData` file
in the local copy of the repository for use by subsequent scripts. 

```{r save-data}
save(SData, TAData, file = here::here("data/SAW_Paper_Data.RData"))
```

\FloatBarrier

# References
Winke, P., Zhang, X., & Pierce, S. J. (2022). 
*A closer look at a marginalized test method: Self-assessment as a measure of speaking proficiency* 
[Manuscript accepted for publication]. Department of Linguistics, Languages, & 
Cultures, Michigan State University. 

Winke, P., Pierce, S. J., & Zhang, X. (2018, October 12-13). 
*Self-assessment works! Continuation-ratio models for testing course and OPIc score effects on oral proficiency self-assessments* 
[Paper presentation]. East Coast Organization of Language Testers 2018 
Conference, Princeton, NJ, United States. 

\FloatBarrier

# Software Information
We use R Markdown to enhance reproducibility. Knitting the source R Markdown 
script generates the PDF file containing explanatory text, R code, and R output.

- We used [RStudio](www.rstudio.org) `r rstudioapi::versionInfo()$version` to 
  work with R and R markdown files. 
- Our software chain looks like this:
  **Rmd file > RStudio > R > rmarkdown > knitr > pandoc > TinyTeX > PDF file**.
- Source script: *`r knitr:::current_input()`*
- Output file: *`r sub(".Rmd", ".pdf", knitr:::current_input(dir = FALSE))`*
- We recommend using [TinyTeX](https://yihui.org/tinytex/) to compile LaTeX 
  files into PDF files. However, it should be viable to use 
  [MiKTeX](https://miktex.org) instead. 
- We used [pandoc](https://pandoc.org) `r pandoc_version()` for this 
  document. 

This document was generated using the following computational environment and 
dependencies: 

``` {r show_citations}
# Check and report whether we used TinyTex or other LaTeX software. 
which_latex()

# Get R and R package version numbers in use.
devtools::session_info()
```

The current Git commit details and status are:

```{r git_details, echo = TRUE, eval=TRUE}
git_report()
```
